---
title: "<center> OMSBA 5300 Data Translation Project </center>"
author: "<center> Elizabeth, John, Julius, Raymond, Schyuler </center> "
date: "<center> 3/21/2021 </center> "
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
      smooth_scroll: true
    number_sections: false
    df_print: paged 
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)
  lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), 
         detach, character.only = TRUE, unload = TRUE)
options(scipen = 999)
```

\

# Purpose of Analysis

- To determine how COVID has affected the health of the retail industry, as measured by employment.
- To compare how COVID has affected the retail industry relative to other industries.
- To figure out who is still working, earning money and has money to spend during the COVID pandemic.

\

## Process of Analysis

- Explore data related to retail, employment, and COVID.
- Perform regression analysis on data.
- Test models for fit and significance.
- Deliver analysis.

\

```{r, message= FALSE, warning=FALSE, echo=FALSE, include=FALSE}
# Libraries
library(car)
library(zoo) 
library(readr)
library(jtools)
library(vtable)
library(ggpubr)
library(ipumsr)
library(ggplot2)
library(viridis)
library(cowplot)
library(janitor)
library(estimatr)
library(gridExtra)
library(lubridate)
library(tidyverse)
library(fastDummies)
```

```{r include=FALSE}
# Common team for analysis
theme_clean <- function(x) {
  theme(panel.border  = element_blank(),
        axis.ticks.y  = element_blank(),
        axis.ticks.x  = element_blank(),
        axis.line     = element_line(color = "gray"),
        axis.text.y   = element_text(color = "#595959", size = 8),
        axis.text.x   = element_text(color = "#595959", size = 6.5),
        plot.title    = element_text(color = "gray29",  size = 11, 
                                     face  = "plain"),
        legend.text   = element_text(color = "#595959", size = 10, 
                                     face  = "plain"),        
        legend.title  = element_text(color = "#595959", size = 10, 
                                     face  = "plain"),
        axis.title    = element_text(color = "#595959", size = 10, 
                                     face  = "plain"),
        plot.subtitle = element_text(color = "gray29",  size = 9, 
                                     face  = "plain"),
        panel.background = element_blank(),
        strip.background = element_rect(colour = "White", fill = "White"))
}
percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Set default working directory
setwd('/Users/cray/Desktop/OMSBA5300/OMSBA5300_DTP/Data')

# Employment data
ddi <- read_ipums_ddi("cps_00006.xml")
data <- read_ipums_micro(ddi)

# State code abbreviation
StateCode <- read_csv('State_FIPS.csv')
StateCode <- StateCode %>% 
  select(State, STATEFIP)

# Industry codes to merge with IPUMS
Industries <- read_csv('indnames.csv')
Industries <- Industries %>%
  mutate(Industry = indname, IND = ind) %>%
  select(-indname, -ind) %>% 
  # Selecting only industries of intrest
  filter(Industry %in% c('Manufacturing', 
                         'Retail Trade', 
                         'Information',
                         'Arts, Entertainment, and Recreation, and Accommodation and Food Services')) %>% 
  mutate(Industry = case_when(Industry == 'Retail Trade' ~ 'Retail',
                              Industry == 'Manufacturing' ~ 'Manufacturing',
                              Industry == 'Information' ~ 'Information',
                              Industry == 'Arts, Entertainment, and Recreation, and Accommodation and Food Services' ~ 'Arts, Leisure and Hospitality'))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Tidy Data

# Removing Haven labels
class_converted <- data %>%
  mutate(Gender = as_factor(lbl_clean(SEX)),
         ASECFlag = as_factor(lbl_clean(ASECFLAG)),
         LaborForce = as_factor(lbl_clean(LABFORCE)),
         FamilyIncome = as_factor(lbl_clean(FAMINC)),
         COVIDTelework = as_factor(lbl_clean(COVIDTELEW)),
         COVIDUnableToWork = as_factor(lbl_clean(COVIDUNAW)),
         FoodStampRecipient = as_factor(lbl_clean(FOODSTMP)),
         COVIDUnableToLookForWork = as_factor(lbl_clean(COVIDLOOK))) %>% 
  select(-c(ASECFLAG, FOODSTMP, SEX, COVIDTELEW, 
            COVIDUNAW, COVIDLOOK, FAMINC, LABFORCE))

# Categorize variables
df <- class_converted %>% 
  mutate(Married = ifelse(MARST == 1, 1, 0)) %>% 
  mutate(Race = case_when(
    HISPAN == 0 & RACE == 100 ~ 'White',
    HISPAN == 0 & RACE == 200 ~ 'Black',
    HISPAN == 0 & RACE %in% c(650, 651, 652) ~ 'Asian',
    HISPAN == 0 & !(RACE %in% c(100, 200, 650, 651, 652)) ~ 'MixedRace',
    HISPAN != 0 ~ 'Latino')) %>% 
  mutate(Regions = case_when(
    REGION == 11 | REGION == 12 ~ 'Northeast',
    REGION == 21 | REGION == 22 ~ 'Midwest',
    REGION == 31 | REGION == 32 | REGION == 33 ~ 'South',
    REGION == 41 | REGION == 42 ~ 'West',
    REGION == 97 ~ 'Unknown')) %>% 
  mutate(AgeGroup = case_when(
    AGE < 20 ~ 'under20', 
    AGE >= 20 & AGE < 30 ~ '20to29',
    AGE >= 30 & AGE < 40 ~ '30to39',
    AGE >= 40 & AGE < 50 ~ '40to49',
    AGE >= 50 & AGE < 60 ~ '50to59',
    AGE >= 60 ~ 'over60')) %>%   
  mutate(MetroArea = case_when(
    METRO %in% c(0, 4, 9) ~ 'Others',
    METRO == 1 ~ 'Rural',
    METRO == 2 ~ 'Urban',
    METRO == 3 ~ 'Suburb')) %>%
  select(-ASIAN, -HISPAN, -RACE, -SERIAL, -METRO, -REGION, -AGE, -MARST) %>%
  
  # Merging data 
  left_join(StateCode, by = 'STATEFIP') %>% 
  select(-STATEFIP) %>% 
  
  inner_join(Industries, by = 'IND') %>% 
  mutate(Employed = case_when(
    EMPSTAT <=  1 | EMPSTAT >= 30 ~ 'Others',
    EMPSTAT >= 10 &  EMPSTAT < 20 ~ '1',
    EMPSTAT >= 20 &  EMPSTAT < 30 ~ '0')) %>%   
  mutate(Education = case_when(
    EDUC >= 10 & EDUC <= 72 ~ 'NoHighschool',
    EDUC == 73 ~ 'HighschoolNoCollege',
    EDUC %in% c(80,81,90,100,110,120,121,122,91,92)
    ~ 'SomeCollegeAssociateDegree',
    EDUC == 111 ~ 'BachelorDegree',
    EDUC %in% c(123,124,125) ~ 'GraduateDegreeOrHigher',
    EDUC %in% c(0, 1, 2, 999) ~ 'NA')) %>%  
  mutate(WorkStatus = case_when(
    WKSTAT %in% c(10,11,14,15) ~ 'Full Time',
    WKSTAT %in% c(12,20,21,22,40,41) ~ 'Part Time',
    WKSTAT %in% c(13,42,50,60) ~ 'Unemployed',
    WKSTAT == 99 ~ 'NA' )) %>% 
  mutate(PartTimeReason = case_when(
    !(WHYPTLWK %in% c(52, 60, 101, 121, 122)) ~ 'Others',
    WHYPTLWK ==  52 ~ 'Job Terminated',
    WHYPTLWK ==  60 ~ 'Could Only Find PT',
    WHYPTLWK == 101 ~ 'Health Limitation',
    WHYPTLWK == 121 ~ 'Childcare Problem',
    WHYPTLWK == 122 ~ 'Other Personal Obligations')) %>%  
  mutate(UnemploymentReason = ifelse(WHYUNEMP==1, 'Layoff', 'Other')) %>% 
  filter(Education != 'NA') %>% 
  filter(Employed  != 'Others') %>% 
  mutate(Employed = as.numeric(Employed)) %>% 
  filter(!is.na(WTFINL)) %>% 
  select(-WKSTAT, -IND, -EDUC, -EMPSTAT, -WHYUNEMP, -WHYPTLWK, -OCC)
```

# The Data

- Data was provided by IPUMS' Current Population Survey.
- Employment statistics were chosen as an indicator of the health of an industry since revenues of the businesses can be misrepresented, or might be affected by government aid.
- Time line: from January 1, 2019 to December 31, 2020
  - March 2020 rows with NAs in WTFINL were eliminated. However, there are still significant amount of data left for the month.
- Industries: 
  - The industries are picked based on the following categories:
    - Essential (Retail) vs. Non-Essential (Arts, Leisure and Hospitality)
    - Remote (Information) vs. In-Person (Manufacturing).
    
\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Show example of data
head(df) 
```

\

# Research Question 1:
**How has COVID affected the health of the retail industry?**

\

## Data Exploration

\

**Retail Data Set**

To determine how COVID has affected the health of the 'Retail Industry', as measured by employment, we filtered the main data set create a retail-specific data set. Our variables of interest that we thought would capture this include variables regarding work during COVID, and status of employment. Variables regarding work during COVID give us a clear look at how the retail industry was specifically impacted by the pandemic. The geography variables were chosen to compare if COVID effected the 'Retail Industry' differently in different communities. All variables were converted to factors. The weighted variable was selected to ensure equal representation from the sample size. We also converted the factor variables 'YEAR' and 'MONTH' to create a continuous 'Date' variable. Finally, we added a variable that grouped months by seasons, to capture the impact of seasonal differences throughout the year on the 'Retail Industry'.

\

```{r message=FALSE, warning=FALSE, include=FALSE}
# Create a data frame for retail
Retail <- df %>%
  select(YEAR, MONTH, Industry, Regions, MetroArea, Employed, LaborForce,
         WorkStatus, COVIDTelework, UnemploymentReason, COVIDUnableToWork, 
         COVIDUnableToLookForWork, PartTimeReason, WTFINL) %>%
  filter(Industry   == 'Retail') %>% 
  mutate(LaborForce =  case_when(
         LaborForce == 'NIU' ~ 'NIU',
         LaborForce == 'No, not in the labor force' ~ 'No',
         LaborForce == 'Yes, in the labor force' ~ 'Yes')) %>% 
  mutate(Employed = as.factor(Employed),
         Urban    = ifelse(MetroArea == "Urban", "Urban", "Not Urban")) %>% 
  mutate_if(is.character, as.factor) 

# Create Day variable, combine with YEAR and MONTH for continuous Date variable
Retail$Date  <- as.yearmon(paste(Retail$YEAR, Retail$MONTH), "%Y %m")
Retail <- transform(Retail, Date = as.Date(Date, frac = 1))
Retail <- Retail %>% 
  mutate(Day = 1, Year = as.numeric(YEAR), MONTHn = as.numeric(MONTH))
Retail$Date <- as.Date(with(Retail, 
                            paste(Year, MONTHn, Day, sep="-")), "%Y-%m-%d")

# Create a variable for season, since the retail industry is often seasonal
Retail <- Retail%>% 
  mutate(Season = case_when(
    MONTHn ==  1 | MONTHn ==  2 | MONTHn ==  3 ~ 'Winter',
    MONTHn ==  4 | MONTHn ==  5 | MONTHn ==  6 ~ 'Spring',
    MONTHn ==  7 | MONTHn ==  8 | MONTHn ==  9 ~ 'Summer',
    MONTHn == 10 | MONTHn == 11 | MONTHn == 12 ~ 'Fall'))
```

\

**Descriptive statistics for variables of interest:**

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# COVID Variables
summary(Retail %>%
          select(COVIDTelework, COVIDUnableToWork, COVIDUnableToLookForWork)) %>%
  kable() %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')
```

\

- COVID variables have 108,264 NAs each (~60% of observations), since these variables only have data available after April 2020. 'COVIDUnableToLookForWork' only has NIU or NA values, so we will exclude this from our analysis. 'COVIDUnableToWork' will be the best variable to use, since it has the most observations with values.

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Employment Variables
summary(Retail %>%
          select(Employed, LaborForce, WorkStatus,
                 UnemploymentReason, PartTimeReason)) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')
```

\

- From the survey, 94% of respondents are employed and 98% are in 'LaborForce'. Only 2% reported layoff as reason for unemployment. 'Employed' numbers are very similar to the numbers in 'LaborForce'. Majority of observations also have full-time employment, but approximately 14.5% of the observations are unemployed. The vast majority of unemployment and part time reasons are marked as 'Other', which means they may not be the best variables to use for understanding employment impacts. 

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Geography Variables
summary(Retail %>% 
          select(MetroArea, Regions)) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')
```

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# geography variables
grid.arrange(
Retail %>% 
  count(MetroArea) %>% 
  ggplot(aes(x = reorder(MetroArea, n, sum), y = n, fill = MetroArea)) + 
  geom_col() +
  theme(axis.ticks  = element_blank(), legend.title = element_blank(), 
        legend.position = 'none') +
  theme_clean() +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", 
                             "#03396c", "#011f4b")) +
  labs(y = 'Population', x = 'MetroAreas'),

Retail %>% 
  count(Regions) %>% 
  ggplot(aes(x = reorder(Regions, n, sum), y = n, fill = Regions)) + 
  geom_col() +
  theme(axis.title.y = element_blank(), axis.ticks = element_blank(), 
        legend.position = 'none') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", 
                             "#03396c", "#011f4b")) +
  theme_clean() +
  labs(x = 'Regions'),
top = 'Geography', ncol = 2)
```

\

- Bar charts give more clarity around the distribution of the population across 'MetroArea' type and 'Regions'. Majority of observations are in the 'South' and 'West', with the least number of observations in the 'Northeast'. For 'MetroArea', 'Suburb' has the highest number of observations, followed by urban areas. However, it's important to note that the majority of observations are in non-urban areas.

\

**Multivariate Exploratory Analysis**

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Average unemployment rates by Year and Season
y <- Retail %>% 
  group_by(Year) %>% 
  summarise(`Average Unemployment` = weighted.mean(Employed == 0, w = WTFINL)) %>%
  adorn_pct_formatting() 

s <- Retail %>% 
  group_by(YEAR, Season) %>% 
  summarise(`Average Unemployment` = weighted.mean(Employed == 0, w = WTFINL)) %>%
  pivot_wider(names_from = YEAR, values_from = `Average Unemployment`) %>% 
  adorn_pct_formatting() %>% 
  as.data.frame()

ys <- cbind(y, s)
ys <- kable(ys) %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')  
ys

```

\

- Overall, the average unemployment rate for the 'Retail Industry' showed a 4.4 percentage point increase between 2019 and 2020. When assessing overall changes to average unemployment, 'Spring' showed the highest overall increase between 2019 and 2020. Furthermore, we see that overall average unemployment decreased in 'Summer' 2020 and Fall 2020, but was still higher than in 2019.

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# Time line from January 2019 through December 2020
Retail %>%
  group_by(Date) %>%
  summarize(UnemploymentRate = 
              1 - (weighted.mean(Employed =='1', w = WTFINL))) %>%
  ggplot(aes(x = Date, y = UnemploymentRate)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_point() +
  geom_line(color = 'blue') +
  theme(axis.title.y = element_blank()) +
  theme_clean() +
  labs(title = 'Average Unemployment Rate for Retail: 2019 to 2020')
```

\

- We see a clear spike in the average unemployment rate at the beginning of 2020, specifically in April 2020. It steadily decreases as we move through the rest of 2020, but the year ends with an overall average unemployment higher than the same time period in 2019. While average unemployment is not as high at the end of 2020 when compared with the beginning of the year, the retail industry is still experiencing a negative impact as unemployment levels are still higher when compared with the end of 2019.

\

**The specific average unemployment figures are shown in the tables below:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Difference in unemployment rate from 2019 to 2020
year19 <- Retail %>% 
  group_by(Date) %>% 
  filter(Date >= '2018-12-30' & Date <= '2019-12-31') %>%  
  summarise(UnemploymentRate19 = 
              1 - (weighted.mean(Employed ==1, w = WTFINL))) %>% 
  adorn_rounding(2)
  
year20 <- Retail %>% 
  group_by(Date) %>% 
  filter(Date >= '2019-12-30') %>%   
  summarise(UnemploymentRate20 = 
              1 - (weighted.mean(Employed ==1, w = WTFINL)))  %>% 
  adorn_rounding(2)

UnemploymentRateDiff <- (year20$UnemploymentRate20 - year19$UnemploymentRate19)
diff <- cbind(year19, year20, UnemploymentRateDiff)
diff2 <- diff %>% 
  adorn_pct_formatting() %>% 
  as.data.frame() 

kable(diff2) %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')
```

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
Months <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
diff$Months <- Months
difff <- diff %>% select(Months, UnemploymentRateDiff)

difff %>%
  select(Months, UnemploymentRateDiff) %>% 
  ggplot(aes(x = Months, y = UnemploymentRateDiff)) +
  theme(legend.position = "none", axis.title.y = element_blank(),
        plot.margin=unit(c(.5,1,.5,.75),"cm"), 
        plot.title = element_text(hjust = 0.25)) +
  geom_point() +
  theme_clean() +
  scale_x_discrete(limits = c(1:12)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_line(color = 'blue') +
  labs(title = 'Difference in Average Unemployment Rate of 2019 vs. 2020')
```

\

- There is little variance for unemployment rates in 2019, with the first couple of months were the same until April. Although COVID shutdown started in March of 2020, unemployment rates did not increase until April 2020, which over 13% higher than April of 2019. However, unemployment rates began to drop gradually each month there after.

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
# COVID Variables
tw1 <- Retail %>%
  group_by(Date) %>%
  filter(Date > '2020-03-31') %>%
  summarise(`Remote Work` = 
              weighted.mean(COVIDTelework == "Yes", w = WTFINL)) %>%
  adorn_pct_formatting(digits = 2) %>% 
  as.data.frame()

uw2 <- Retail %>%
  group_by(Date) %>%
  filter(Date > '2020-03-31') %>%
  summarise(`Unable To Work` = 
              weighted.mean(COVIDUnableToWork == "Yes", w = WTFINL)) %>%
  adorn_pct_formatting(digits = 2) %>% 
  as.data.frame()

tw <- cbind(tw1, uw2)
kable(tw) %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered', full_width = FALSE) %>% 
  row_spec(0, background = 'lightblue')
```

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
te <- Retail %>%
  group_by(Date) %>%
  filter(Date > '2020-03-31') %>%
  summarise(`Remote Work` = 
              weighted.mean(COVIDTelework == "Yes", w = WTFINL)) %>% 
  ggplot(mapping = aes(x = Date, y = `Remote Work`)) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  geom_line(color = 'blue') +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  theme_clean() +
  labs(title = 'Pct. of Retail Industry Teleworking due to COVID')

un <- Retail %>%
  group_by(Date) %>%
  filter(Date > '2020-03-31') %>%
  summarise(`Unable To Work` = 
              weighted.mean(COVIDUnableToWork == "Yes", w = WTFINL)) %>% 
  ggplot(mapping = aes(x = Date, y = `Unable To Work`)) +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_line(color = 'blue') +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme_clean() +
  labs(title = 'Pct. of Retail Industry Unable to Work due to COVID')

grid.arrange(te, un, nrow = 1)
```

\

- COVID related data was not provided until May of 2020. It is also important to note that these variables have a high amount of NA and NIU values. However, we are using them because we think they capture the unique employment impacts to the 'Retail Industry' due to the pandemic.

- Only about 14% of the 'Retail Industry' was able to work remotely at the start of the pandemic under 9% were working remotely by the end of 2020. In May 2020, about 29% responded that they were unable to work. However, this steadily decreased throughout the year and by December 2020 5.4% reported that they were unable to work. These trends match the overall unemployment trends for 2020.

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# Graph the average unemployment by region, comparing 2019 and 2020
Retail %>% 
  mutate(YEAR = as.factor(YEAR)) %>% 
  group_by(Regions, YEAR) %>% 
  summarise(AverageUnemploymentRate = 
              weighted.mean(Employed == 0, w = WTFINL)) %>%
  ggplot(mapping = aes(x = Regions, y = 
                         AverageUnemploymentRate, fill = YEAR)) +
  theme(axis.title.y = element_blank()) +
  theme_clean() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_bar(position = 'dodge', stat = 'identity') +
  scale_fill_manual(values=c( "#6497b1", "#03396c")) +
  ggtitle('Average Unemployment Rate by Region: 2019 vs. 2020') 
```

\

- All 'Regions' show higher average overall unemployment rates in 2020 than in 2019, with the 'Northeast' showing the highest overall increase in average unemployment. The south shows the lowest overall increase in the average unemployment rate between 2019 and 2020 while the 'Northeast' has the highest overall unemployment rate in 2020 and the highest percentage point increase between 2019 and 2020.

- What could be causing this difference? Are there differences in 'MetroArea' types between the 'Regions'? Do some metro areas show a more substantial increase in average unemployment rates than others?

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# Graph the average unemployment over time, comparing across regions
Retail %>% 
  group_by(Regions, Date) %>% 
  summarise(AverageUnemploymentRate = 
              weighted.mean(Employed == 0, w = WTFINL)) %>%
  ggplot(mapping = aes(x = Date, y = AverageUnemploymentRate, 
                       color = Regions)) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  theme_clean() +
  ggtitle('Average Unemployment Rate by Region: 2019 vs. 2020') 
```

\

- All 'Regions' follow the same pattern of low average unemployment rates in 2019, followed by a significant spike in early 2020 and a steady downward trend for the rest of 2020. All 'Regions' show higher average unemployment rates at the end of 2020 compared with 2019.

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
Retail %>%
  group_by(Date, Regions) %>%
  filter(Date > '2020-03-31') %>%
  summarise(`Remote Work` = 
              weighted.mean(COVIDTelework == "Yes", w = WTFINL)) %>%
  ggplot(mapping = aes(x = Date, y = `Remote Work`, color = Regions)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank() )+
  theme_clean() +
  ggtitle('Proportion of Retail Industry Working Remotely by Region 2020')
```

\

- The 'West' started the year with the highest proportion of retail employees working remotely and is overall one of the regions with the highest proportions of remote work. The 'South' overall has the lowest proportion of remote workers, and this region decreased most sharply over the course of the year. 'West' and 'Northeast' had an increase toward the end of the year, 'South' and 'Midwest' showed decreases. Additionally, not every region follows the same trend as the 'Retail Industry' overall. While the overall industry did see a slight increase at the end of the year, some 'Regions' continued to decrease while others increased.

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# Unable to work by region
Retail %>%
  group_by(Date, Regions) %>%
  filter(Date > '2020-04-30') %>%
  summarise(`Unable to Work` = weighted.mean(COVIDUnableToWork == "Yes", w = WTFINL)) %>%
  ggplot(mapping = aes(x = Date, y = `Unable to Work`, color = Regions)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank() )+
  theme_clean() +
  ggtitle('Proportion of Retail Industry Unable to Work by Region 2020')
```

\

- Overall, all 'Regions' followed the same pattern with regard to the percentage of retail workers unable to find work in 2020. Around 30% of retail workers in each region was unable to find work in April 2020, but then every region showed a sharp decrease soon after, which started to slow down after July 2020. However, the 'South' had the lowest percentage while the 'West' and 'Northeast' consistently had the highest 'Regions'.

\

```{r echo=FALSE, fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
# Unemployment by Regions and MetroArea the month before and after COVID outbreak
Retail %>%
  mutate(MONTH = as.factor(MONTH)) %>%
  group_by(MetroArea, MONTH, Regions) %>%
  filter(MONTH == '2' | MONTH == '4') %>%
  summarize(UnemploymentRate =
              1 - (weighted.mean(Employed =='1', w = WTFINL))) %>%
  ggplot(mapping = aes(x = MetroArea,
                       y = UnemploymentRate, fill = MONTH)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_bar(position = 'dodge', stat = 'identity') +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank()) +
  scale_fill_discrete(name = "2020", labels = c('Feb','April')) +
  guides(fill = guide_legend(reverse = TRUE)) +
  ggtitle('Average Unemployment Rate') +
  scale_fill_manual(values=c( "#6497b1", "#03396c")) +
  theme_clean() +
  facet_grid(Regions ~ .)
```

\

- All 'Regions' have higher unemployment rates in 2020 when compared with 2019. However, 'Rural' areas showed the smallest overall increase when comparing the two years. 'Urban' and 'Suburb' areas show the highest overall increases. 
- Are there any notable differences when looking at 'MetroAreas' in 'Regions'?

- Some regions have more consistent average unemployment than others. The 'MetroArea' in the 'South' have around the same percentage of average unemployment, while other 'Regions' show greater differences between their 'MetroArea'. The 'Northeast' especially shows some large differences when comparing 'Urban' and 'Suburban' 'MetroArea' to 'Rural' and 'Others'.

\

```{r echo=FALSE, fig.height=3, fig.width=9, message=FALSE, warning=FALSE}
# Compare unemployment rates across regions and metro areas for 2020
Retail %>% 
  filter(YEAR == 2020) %>% 
  mutate(YEAR = as.factor(YEAR)) %>% 
  group_by(MetroArea, Regions, YEAR) %>% 
  summarise(AverageUnemployment = weighted.mean(Employed == 0, w = WTFINL)) %>% 
  ggplot(mapping = aes(x = MetroArea, y = AverageUnemployment, fill = Regions)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  geom_bar(position = 'dodge', stat = 'identity') +
  theme(axis.title.y = element_blank())+
  ggtitle('Average Unemployment in 2020 for Region Within Metro Area') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96",
                             "#03396c", "#011f4b")) +  
  theme_clean()
```

\

- There appears to be notable differences when comparing average unemployment rates by 'MetroArea' across 'Regions'. Overall, 'Rural' appears to have the smallest overall increase in unemployment rate in all 'Regions', while 'Urban' appears to have the largest. 'Urban' areas in the 'Northeast' have stayed consistently the highest average unemployment rates in 2020.

\

```{r echo=FALSE, fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
# Difference in unemployment rate 1 month before/after forRegions/MetroAreas
feb <- Retail %>%
  group_by(MONTH, Regions, MetroArea) %>%
  filter(MONTH == 2) %>%
  summarize(UnemploymentRate =
              1 - (weighted.mean(Employed == 1, w = WTFINL))) %>%
  adorn_rounding(2)

may <- Retail %>%
  group_by(MONTH, Regions, MetroArea) %>%
  filter(MONTH == 4) %>%
  summarize(UnemploymentRate =
              1 - (weighted.mean(Employed == 1, w = WTFINL))) %>%
  adorn_rounding(2)

fm <- merge(feb, may, by = c("Regions", "MetroArea"))
fm <- fm %>%
  mutate(UnemploymentRate_Diff = fm[6] - fm[4]) %>%
  select(-MONTH.x, -MONTH.y) %>% 
  as.data.frame() %>% 
  adorn_pct_formatting()

names(fm)[3] <- 'FebruaryUnemploymentRate'
names(fm)[4] <- 'AprilUnemploymentRate'

kable(fm) %>% 
  kable_styling(bootstrap_options = 'striped', 'bordered') %>% 
  row_spec(0, background = 'lightblue')
```

\

- My assumption was that densely populated areas like the 'Northeast' 'Regions' and 'Urban' areas would be hit the hardest. However, with the weighted means, it shows that it was the 'Midwest', 'Urban' and 'Suburb' that showed the biggest difference in unemployment.

\

## Models

- The first model, looking for the effect of COVID on employment in the 'Retail' industry by 'Regions'.

\

$$ Employed= \beta_0 + \beta_1Regions_i + \beta_2MetroArea_i $$
\

- The second model, looking for the effect of COVID on employment in the 'Retail' industry by 'MetroArea', controling for 'MONTH'.

\

$$ Employed= \beta_0 + \beta_1Regions_i + \beta_2MetroArea_i + \beta_3MONTH_i + \beta_4COVIDUnableToWork_i  $$

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
Retail$Employed <- as.numeric(Retail$Employed)

Model1 <- lm_robust(Employed ~ Regions + MetroArea, data = Retail, weights = WTFINL)

Model2 <- lm_robust(Employed ~ Regions + MetroArea + MONTH + COVIDUnableToWork, 
                    data = Retail %>% mutate(MONTH = as.factor(MONTH)), 
                    weights = WTFINL)
```

\

### Model Results:

\

```{r echo=FALSE, message=FALSE, warning=FALSE}
export_summs(Model1, Model2, 
             model.names = c('Regions and MetroArea', 'Month and COVID'))
```

\

**Key Findings**

*Regions and Metro Area*

Overall, most regions and metro areas did not have an impact on the probability of being employed in the retail industry. The two exceptions were the Northeast region and Urban areas, both of which were significant at the 95% level of confidence. We interpret their coefficients as follows:

- Holding all else constant, there is a 2 percentage point lower probability of being employed in the Northeast as opposed to the reference region Midwest, within the reference metro area of Other.

- Holding all else constant, there is a 2 percentage point lower probability of being employed in Urban areas as opposed to the reference area of Other areas, within the reference region of Midwest.

\

*Month and COVID*

In this model, the coefficients for Northeast and Urban are largely unchanged and remain statistically significant. Being unable to work due to business closures and month of year are also statistically significant in this model. We interpret the coefficients as follows:

- Holding all else constant, there is a 29 percentage point lower probability of being employed if someone answered 'Yes' to being unable to work due to COVID. 

For months, all coefficients are statistically significant at the 95% level of confidence. Additionally, the reference months for all coefficients is May 2020.

- Holding all else constant, there is a 2 percentage point increase in the probability of being employed in June 2020 as opposed to the reference month. 

- Holding all else constant, there is a 2 percentage point increase in the probability of being employed in July 2020 as opposed to the reference month. 

- Holding all else constant, there is a 3 percentage point increase in the probability of being employed in August 2020 as opposed to the reference month. 

- Holding all else constant, there is a 2 percentage point increase in the probability of being employed in September 2020 as opposed to the reference month. 

- Holding all else constant, there is a 3 percentage point increase in the probability of being employed rea of Other.

- Holding all else constant, there is a 3 percentage point increase in the probability of being employed in November 2020 as opposed to the reference month. 

- Holding all else constant, there is a 4 percentage point increase in the probability of being employed in December 2020 as opposed to the reference month. 

\

## Testing

```{r echo=FALSE, message=FALSE, warning=FALSE}
linearHypothesis(Model1, matchCoefs(Model1, "Regions"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
linearHypothesis(Model2, matchCoefs(Model2, "MetroArea"))
```

\

## Summary of Findings  

**How has COVID affected the health of the retail industry?**

\

Overall, the retail industry was impacted severely by the pandemic during 2020. Our models show that if a person in the retail industry was unable to work due to business closures for COVID, that there was a 29 percentage point less probability of being employed.

When looking at the impact throughout 2020, the retail industry saw a larger negative impact at the beginning of 2020 when compared with the last half of the year. Unemployment rates for the retail industry spiked to 18% in April 2020, but began to decrease throughout the year. By December 2020, unemployment was rate was at 6%. Additionally, we saw a comparable decrease throughout 2020 in the percentage of the retail workforce that was unable to work. In May 2020 about 28.8% of the industry was unable to work, which decreased to 5.40% by December 2020. However, while there was huge improvement in unemployment throughout the year, year-end unemployment rates in retail were still higher when compared with those same months in 2019.

Finally, we see the most severe impact in the Northeast and in Urban areas. All regions and metro areas showed higher unemployment rates in 2020 when compared with 2019, and all regions showed a spike in their unemployment rates in April 2020. However, the Northeast was hit the hardest and is associated with an estimated 2 percentage point lower probability of being employed. Urban areas were also hit hard relative to other metro areas, also being associated with an estimated 2 percentage point lower probability of being employed.

In conclusion, while the retail industry was able to make a rebound by years' end, the industry is still in slightly a worse position relative to 2019 due to slightly higher unemployment rates when compared with 2019. We also see that the impact to the retail industry is worse in the Northeast and in Urban areas.
  
\

# Research Question 2:
**How has retail fared relative to other industries?**

\

## Data Exploration

\

**Overall employment figures by industry, year over year**

\

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Explore number of people employed by industry
jobs_industry <- df %>% 
  group_by(Industry, YEAR, Employed) %>% 
  count()
# Convert to factor
jobs_industry$YEAR <- as.numeric(jobs_industry$YEAR)
jobs_industry$Industry <- as.factor(jobs_industry$Industry)
jobs_industry$Employed <- factor(jobs_industry$Employed)
jobs_industry$Employed <- factor(jobs_industry$Employed, levels = c("1", "0"), 
                  labels = c("Yes", "No"))
# Plot jobs by industry
p1 <- ggplot(data = jobs_industry)+
  aes(x = YEAR)+
  aes(y = n) +
  aes(group = Employed)+
  facet_wrap(~ Industry)+
  geom_line(aes(linetype = Employed, color = Employed)) +
  geom_point(aes(color = Employed)) +
  scale_x_continuous(breaks = c(2019, 2020))+
  scale_y_continuous(labels = scales::comma ) +
  scale_color_manual(values=c("#6497b1", "#011f4b")) +
  theme_clean() +
  theme(legend.key=element_blank())+
  labs(title = "Retail and Manufacturing are similarly represented in the data and show similar overall \njob number trends", 
       subtitle = "Arts, Leisure and Hosptility lost the most  total jobs due to COVID19", 
       x = "Year", y = "Total Employed", 
       tag = , caption = "Source: IPUMS CPS Monthly Data")
p1
```

\

- Overall job numbers show a decrease from 2019 to 2020 for all industries. 

- `Manufacturing` and `Retail` are similarly represented in the data and show similar trends. `Arts, Leisure and Hospitality` shows the most dramatic drop in employment numbers

- Albeit less represented in the total sample, the `Information` industry appears to be the less impacted in terms of overall jobs, ear over year

\

**Unemployment figures by industry**

\

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Filter years
df_2019 <- df %>% 
  filter(YEAR == 2019)
df_2020 <- df %>% 
  filter(YEAR == 2020) 
df_2020_COVID <- df %>% 
  filter(YEAR == 2020 & (!MONTH %in% c(1, 2, 3)))
## Unemployment Rate by Industries
unemp_ind_2020 <- df_2020 %>% 
  group_by(Industry, MONTH, Gender, Race, Education) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))
p2 <- ggplot(data = unemp_ind_2020, aes(x = MONTH, y = AvgUnempRate, color = Industry, shape = Industry)) +
  #geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  xlim('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec') +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values=c("darkgray", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme_clean() +
 geom_vline(aes(xintercept = 4 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Apr", y = .03, label = 'COVID Shutdown', vjust= .5, hjust = 1, fill = 'white', size = 7/.pt, color = "gray29" ) +
  geom_vline(aes(xintercept = 7 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Jul", y = .03, label = 'Reopening Paused', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  geom_vline(aes(xintercept = 5 ), linetype = 'dashed', color = 'lightblue3') +
  annotate(geom = 'text', x = "May", y = .06, label = 'Partial Reopening', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  labs(title = "Unemployment rates rose dramatically at the onset of the COVID-19 Pandemic", 
       subtitle = "Pausing the reopening of economy kept retail from improving more quickly", 
       x = "Month", y = "Avg Unemployment Rate", 
       tag = , caption = "Source: IPUMS CPS Monthly Data")
p2
```

\

- `Arts, Leisure and Hospitality` is the only industry more affected than `Retail` in terms of unemployment rates. 

- The partial reopening helped all industries recover from their April unemployment highs, with the pause on the reopening stalling each industry's recovery. When compared to `Retail`, `Manufacturing` has a lower unemployment peak and less of a sustained, higher unemployment rate 

- The `Information` industry drop in unemployment rate really took off in July, outpacing `Retail`recovery by November. 

- In spite of the traditional peak season for the `Retail` industry, unemployment trend does not significantly improve in terms of unemployment rate. 

\

**Unable to work due to COVID19**

\

```{r, echo = FALSE, message=FALSE}
df_2020_UnableToWork <- df_2020 %>% 
  filter(COVIDUnableToWork != 'NIU') %>% 
  mutate(UnableToWorkCOVID = ifelse(COVIDUnableToWork == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)
## COVIDUnableToWork Breakdown by Employed
UnableToWorkInd_2020 <- df_2020_UnableToWork %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgUnableToWork = 1 - weighted.mean(UnableToWorkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))
UnableToWorkInd_2020$Employed <- factor(UnableToWorkInd_2020$Employed)
# Factor industry
UnableToWorkInd_2020$Industry <- factor(UnableToWorkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))
# Plot 
p9 <- ggplot(data = UnableToWorkInd_2020, aes(x = Employed, y = AvgUnableToWork, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgUnableToWork), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Average Rate of Reported Not Able To Work Due to COVID') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  facet_wrap(~ Industry)+
  theme_clean()+
  theme(legend.position="none") +
  labs(title = "Manufacturing reported the highest percentage of being unable to work due to COVID-19, \nfollowed by Retail ", 
       subtitle = "", 
       x = "Month", y = "% Reporting Unable to Work due to COVID-19", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 
df_2020_Telework <- df_2020 %>% 
  filter(COVIDTelework != 'NIU') %>% 
  mutate(TeleworkCOVID = ifelse(COVIDTelework == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)
## COVIDTelework Breakdown by Employed
TeleworkInd_2020 <- df_2020_Telework %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgTelework = 1 - weighted.mean(TeleworkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))
# Factor industry
TeleworkInd_2020$Industry <- factor(TeleworkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))
p10 <- ggplot(data = TeleworkInd_2020, aes(x = Employed, y = AvgTelework, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgTelework ), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  facet_wrap(~ Industry) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  theme(legend.position="none") +
  labs(title = "Persons still employed in the Retail and Arts, Leisure and Hospitality industries \nreport the highest percentages of being able to telework ", 
       subtitle = "", 
       x = "Employed", y = "% Reporting Able to Telework", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 
# Telework
df_2020_Telework <- df_2020 %>% 
  filter(COVIDTelework != 'NIU') %>% 
  mutate(TeleworkCOVID = ifelse(COVIDTelework == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)
## COVIDTelework Breakdown by Employed
TeleworkInd_2020 <- df_2020_Telework %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgTelework = 1 - weighted.mean(TeleworkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))
# Factor telework
TeleworkInd_2020$Industry <- factor(TeleworkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))
p10 <- ggplot(data = TeleworkInd_2020, aes(x = Employed, y = AvgTelework, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgTelework ), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  facet_wrap(~ Industry) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  theme(legend.position="none") +
  labs(title = "Persons still employed in the Retail and Arts, Leisure and Hospitality industries \nreport the highest percentages of being able to telework ", 
       subtitle = "", 
       x = "Employed", y = "% Reporting Able to Telework", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 
```

```{r, echo=FALSE}
# Print plots
p9
```

\

- `Retail` reported the highest percentage of being unable to work due to COVID-19 .

- While `Arts, Leisure and Hospitality` continues to be the most affected industry among the four looked at, the percentage of persons responding 'no' to being unable to work due to the pandemic was lowest in the  `Information` industry. 

\

```{r, echo=FALSE}
# Print plots
p10
```

\

- Persons still employed in the `Retail` and `Arts,Leisure and Hospitality` industries report the highest percentages of being able to telework.

- By comparison, those still employed in `Information`, a higher sub-group than other industries, report the lowest percentage of being able to telework.

\

### Summary of Exploratory Findings

\

In looking at the data, overall job numbers show a decrease from 2019 to 2020 for all industries during 2020 due to the shutdowns caused by the COVID-19 pandemic. Manufacturing and Retail are similarly represented in the data and show similar trends. Three inflection points present themselves in the data:

1)	The States shutting down in March due to COVID-19.
2)	The partial reopening that some States undertook in May.
3)	The pause on reopening that occurred in July.

Since the staggered nationwide shutdowns began in March, we used data from April- December 2020 to get a more accurate effect on industry employment rates. The results show that Arts, Leisure and Hospitality is the only industry more affected than Retail in terms of unemployment rates during the pandemic.

The partial reopening helped industries recover from their April unemployment highs but the July pause in reopening stalled recovery for all analyzed industries. When compared to Retail,  Manufacturing has a lower unemployment peak and a less prolonged elevation in unemployment rates .The Information industry's improvement in unemployment rates gained momentum in July, outpacing Retail’s recovery by November 2020.

When looking at those who were unable to work because their employer closed or lost business due to the coronavirus pandemic, persons employed in `Retail` and `Manufacturing` occupations were among the most likely to have been unable to work due to the pandemic during the period of April-December 2020. 

\

## Models

\

- Based on the findings in the exploratory analysis, we will be looking at the effect of industry type on the probability of being employed. Additionally, we will be looking at what effect resulted on the probability of being employed when we interact both `Month` and `Industry` in Retail and the other industries. We will also measure the effect `COVIDUnableToWork` and `CovidTelework` had on the same. We will be confirming if second to Arts, Leisure and Hospitality, Retail was the second most affected industry by COVID-19. Additionally, due to the seasonal nature of `Retail` we look to confirm our findings that `Months` played a role in the probability of being employed. 

\

- As a baseline, we run a model looking solely at the effect the type of industry has on the probability being employed.

\

$$ Employed_i= \beta_0 + \beta_1Industry_i    $$

\

- Our second model looks at the effect of month on the industries, particularly the months associated with the initial COVID-19 shutdown (April), the second shutdown (July), and the peak `Retail` season
of November and December.

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2Month_i\times Industry_i   $$
\

- The third model measures the effect of the interaction between `Industry` and `COVIDUnableToWork` and its effect on the probability of being employed. 

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2COVIDUnableToWork_i\times Industry_i   $$

\

```{r, echo=FALSE, include=FALSE}
# Base model
Model_1 <- lm_robust(Employed ~ Industry , data = df_2020, weights = WTFINL)
# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_1, matchCoefs(Model_1, "Industry"))
# Factor Month
df_2020$MONTH = factor(df_2020$MONTH)
# Industry * Month
Model_2 <- lm_robust(Employed ~ Industry + Industry*MONTH , data = df_2020, weights = WTFINL)
# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_2, matchCoefs(Model_2, "Industry"))
```

\

```{r, echo=FALSE, include=FALSE}
## COVID MODELS
# Model with interaction Industry * COVIDUNABLE TO WORK
Model_3 <- lm_robust(Employed ~ Industry + Industry*COVIDUnableToWork, data = df_2020_COVID, weights = WTFINL)
# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_3, matchCoefs(Model_3, "Industry"))
# Factor Month
df_2020_Telework$COVIDTelework = factor(df_2020_Telework$COVIDTelework)
```

\

### Model Results:  
**Industry, Industry:Month, Industry:COVIDUnableToWork**

\

```{r, echo=FALSE}
# Compare Models 1, 2
export_summs(Model_1, Model_2, Model_3,
             model.names = c("Industry", "Industry:Month", "Industry:COVIDUnableToWork"),
             coefs = c("Information" = "IndustryInformation",
                       "Manufacturing" = "IndustryManufacturing",
                       "Retail" = "IndustryRetail",
                       "April" = "MONTH4",
                       "July" = "MONTH7",
                       "November" = "MONTH11",
                       "December" = "MONTH12",
                       "Information:April" = "IndustryInformation:MONTH4",
                       "Manufacturing:April" = "IndustryManufacturing:MONTH4",
                       "Retail:April" = "IndustryRetail:MONTH4",
                       "Information:July" = "IndustryInformation:MONTH7",
                       "Manufacturing:July" = "IndustryManufacturing:MONTH7",
                       "Retail:July" = "IndustryRetail:MONTH7",
                       "Information:November" = "IndustryInformation:MONTH11",
                       "Manufacturing:November" = "IndustryManufacturing:MONTH11",
                       "Retail:November" = "IndustryRetail:MONTH11",
                       "Information:December" = "IndustryInformation:MONTH12",
                       "Manufacturing:December" = "IndustryManufacturing:MONTH12",
                       "Retail:December" = "IndustryRetail:MONTH12",
                       "Retail:COVID-AbleToWork" = "IndustryRetail:COVIDUnableToWorkYes",
                       "Manufacturing:COVID-AbleToWork" = "IndustryManufacturing:COVIDUnableToWorkYes",
                       "Information:COVID-AbleToWork" = "IndustryInformation:COVIDUnableToWorkYes"),
                scale = TRUE, robust = TRUE)
```

\

**Key Findings**

\

$$ Employed_i= \beta_0 + \beta_1Industry_i $$

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2Month_i\times Industry_i   $$

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2COVIDUnableToWork_i\times Industry_i   $$

\

- **Industry**

  We found the following to be statistically significant at the 95% confidence level:

  - With everything else held constant, there is a 10 percentage point higher probability of being employed in the 'Retail' industry as opposed to the reference industry, `Arts, Leisure and Entertainment`.

  - With everything else held constant, there is a 12 and 11 percentage point higher probability of being employed (respectively) in the `Information` and `Manufacturing` industries as opposed to the reference industry, `Arts, Leisure and Entertainment.`

- **Industry:Month**

  We found the following to be statistically significant at the 95% confidence level:

  - With everything else held constant, there is a 32 percentage point lower probability of being employed in the month of April 2020 as opposed to the reference month of January 2020, within the reference industry, `Arts, Leisure and Entertainment.`

  - With everything else held constant, there is a 18 percentage point lower probability of being employed in the month of July 2020 as opposed to the reference month of January 2020, within the reference industry, `Arts, Leisure and Entertainment.`

  - With everything else held constant, there is a 9 percentage point lower probability of being employed in the month of November 2020 as opposed to the reference month of January 2020, within the reference industry, `Arts, Leisure and Entertainment.`

  - With everything else held constant, there is a 10 percentage point lower probability of being employed in the month of December 2020 as opposed to the reference month of January 2020, within the reference industry, `Arts, Leisure and Entertainment.`

  - With everything else held constant, there is a 19 percentage point higher probability of being employed in the month of April 2020 for the `Retail` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.
  
  - With everything else held constant, there is a 22 percentage point higher probability of being employed in the month of April 2020 for either the `Information` or `Manufacturing` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 13 percentage point higher probability of being employed in the month of July 2020 for either the `Retail` or `Manufacturing` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 8 percentage point higher probability of being employed in the month of July 2020 for the `Information` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 8 percentage point higher probability of being employed in the month of November 2020 for the `Retail` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 3 percentage point higher probability of being employed in the month of November 2020 for the `Information` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 7 percentage point higher probability of being employed in the month of November 2020 for the `Manufacturing` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 9 percentage point higher probability of being employed in the month of December 2020 for either the `Retail` or `Manufacturing` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.

  - With everything else held constant, there is a 6 percentage point higher probability of being employed in the month of December 2020 for the `Information` industry as opposed to the month of January 2020 for the `Arts, Leisure and Entertainment` industry.
  
- **Industry:Covid-AbleToWork**

  We found the following to be statistically significant at the 95% confidence level:

  - With everything else held constant, people being able to work during the COVID-19 pandemic in the `Retail` industry have a 6 percentage point higher probability of staying employed as opposed to people who are unable to work during the pandemic in the reference industry,  `Arts, Leisure and Entertainment`.
  
  - With everything else held constant, people being able to work during the COVID-19 pandemic in the `Manufacturing` industry have a 4 percentage point higher probability of staying employed as opposed to people who are unable to work during the pandemic in the reference industry,  `Arts, Leisure and Entertainment`.
  
  - With everything else held constant, people being able to work during the COVID-19 pandemic in the `Information` industry have a 1 percentage points higher probability of staying employed as opposed to people who are unable to work during the pandemic in the reference industry,  `Arts, Leisure and Entertainment`.

\

## Summary of Findings

While other variables held constant, when we compare Retail to Arts, Leisure and Entertainment, there is a 10 percentage points higher probability of being employed in 2020, while we are in the COVID-19 pandemic. Compared to Retail, being in either the Manufacturing or Information industries has a higher probability of being employed by 2 and 1 percentage points, respectively. 

Our exploratory analysis revealed that April and July were two critical months in 2020 that pointed at increased unemployment rates due to either States shutting down or pausing their reopening. The analysis confirms this. The probability of being employed decreases by 32 percentage points in April and by 18 percentage points in July, within the reference industry, Arts, Leisure and Entertainment. In April alone, when comparing to Retail, there is a 3 percentage points higher probability of being employed in both the Manufacturing and Information industries.

Interestingly, when States paused reopening plans in July, being employed in Retail results in a 5-percentage point higher probability of being employed when compared to the Information industry and a 13-percentage point higher probability when compared to Arts, Leisure and Hospitality. 

As November and December are considered peak season for Retail, we compared its performance to the other industries for those months. When compared to the Information industry, Retail enjoys a 5-percentage point advantage in the probability of being employed in November and a 3-percentage point advantage in December. When compared to Arts, Leisure and Hospitality, the same December probability advantage for Retail increases to 9 percentage points. 

Finally, looking at the special COVID-19 question of  "At any time in the last 4 weeks, were you unable to work because your employer closed or lost business due to the coronavirus pandemic?" reveals that answering "Yes" to this question and being in the Retail industry has a 5 percentage points higher probability of being employed when compared to the Information industry and a 6 percentage points advantage over the probability of being employed in the Arts, Leisure and Hospitality industry. 

\

# Research Question 3: 
**Who has money to spend?**

\

## Data Exploration  
 
\

**Overview of employment by Race, Gender, and Education from January 2019 to December 2020** 

\

**By Race** 

\

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Where are the different race working?
demographics <- df %>% 
  group_by(Race = as_factor(Race),
           Industry = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
# demographics

# Plot graph
plot1 <- ggplot(data = demographics, aes(x = Race, reorder(Race), y = Percent, fill = Industry)) +
  geom_bar(position="stack", stat="identity")+
  scale_y_continuous(labels = scales::percent) +
  ylab('Avg Employment Rate') +
  xlab('Race') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  labs(title = "Industries among different races (2019-2020)", 
       subtitle = "Within each race, over 59% are employed by the Retail and Arts, Leisure and Hospitality industries", 
       x = "Race", y = "Avg Percent of Total Employment", 
       tag = , caption = "")
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Filter years
df_2019 <- df %>% 
  filter(YEAR == 2019)

df_2020 <- df %>% 
  filter(YEAR == 2020)

# 2020 unemploymnet among the different races
unemp_ind_2020 <- df_2020 %>% 
  group_by(MONTH, Race) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))

# Plot graph
plot2 <- ggplot(data = unemp_ind_2020, aes(x = MONTH, y = AvgUnempRate, color = Race)) +
  #geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  xlim('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec') +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values=c("#b3cde0","darkgray", "#6497b1", "#005b96", "#03396c" ))+
  theme_clean() +
  geom_vline(aes(xintercept = 4 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Apr", y = .03, label = 'COVID Shutdown', vjust= .5, hjust = 1, fill = 'white', size = 7/.pt, color = "gray29" ) +
  geom_vline(aes(xintercept = 7 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Jul", y = .03, label = 'Reopening Paused', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  geom_vline(aes(xintercept = 5 ), linetype = 'dashed', color = 'lightblue3') +
  annotate(geom = 'text', x = "May", y = .06, label = 'Partial Reopening', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  labs(title = "Black Americans experienced a larger increase in unemployment during \nCOVID-19 shutdown", 
       subtitle = "Asian and White Americans had the most steady recovery in unemployment among all groups", 
       x = "Month", y = "Avg Unemployment Rate", 
       tag = , caption = "")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Show plots
plot1
plot2
```

\

- From 2019 to 2020, the `Retail` and `Arts, Leisure and Hospitality` industries (combined) employed 72.32% of Mixed Race Americans, 68.40% of Latino Americans, and 67.10% of Black Americans.

- Black Americans experienced the most significant increase in unemployment during the COVID-19 shutdown.  The spike in unemployment rates for White Americans was lower compared to other races.

- Asian and White Americans had the most steady recovery in unemployment among all groups, but unemployment recovery is slower for Black Americans.  

\

**By Gender**

\

```{r, echo = FALSE, message=FALSE, warning=FALSE}

# Women's ageGroup & Head of household
demographics <- df %>% 
  filter(PERNUM == 1,
         Gender == "Female") %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(HWTFINL)) %>% 
  mutate(PercentFemaleHeadHousehold = n/sum(n)) 
# demographics

## Industry - Gender
unemp_ind_2020 <- df_2020 %>% 
  group_by(Industry, Gender) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))
# unemp_ind_2020

unemp_ind_2020 <- df_2020 %>% 
  group_by(Industry, MONTH, Gender) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))

# Plot graph
plot3 <- ggplot(data = unemp_ind_2020, aes(x = MONTH, y = AvgUnempRate, color = Gender)) +
  #geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  xlim('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec') +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme_clean() +
   geom_vline(aes(xintercept = 4 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Apr", y = .03, label = 'COVID Shutdown', vjust= .5, hjust = 1, fill = 'white', size = 7/.pt, color = "gray29" ) +
  geom_vline(aes(xintercept = 7 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Jul", y = .03, label = 'Reopening Paused', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  geom_vline(aes(xintercept = 5 ), linetype = 'dashed', color = 'lightblue3') +
  annotate(geom = 'text', x = "May", y = .06, label = 'Partial Reopening', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  labs(title = "Women experienced a higher peak in unemployment during COVID-19 shutdown", 
       subtitle = "Unemployment rates are more sustained among women during Jun-Dec", 
       x = "Month", y = "Avg Unemployment Rate", 
       tag = , caption = "")

# Which indsutry has the highest female employment?
demographics <- df_2020 %>% 
  filter(Gender=="Female") %>% 
  group_by(Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
# demographics

# Which indsutry has the highest men employment?
demographics2 <- df_2020 %>% 
  filter(Gender=="Male") %>% 
  group_by(Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentMale = n/sum(n)) 
# demographics2

# what does Female  & unemployment proportion look like in each industry? 
demographics <- df_2020 %>% 
  group_by(Industries = as_factor(Industry),
           Gender = as_factor(Gender)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
# demographics

## Industry - Gender 
unemp_ind_2020 <- df_2020 %>% 
  group_by(Industry, Gender) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))
# unemp_ind_2020

# Factor Industry
unemp_ind_2020$Industry <- factor(unemp_ind_2020$Industry, levels =c("Manufacturing", "Information", "Retail",  "Arts, Leisure and Hospitality"))

# Plot graph
plot4 <- ggplot(data = unemp_ind_2020, aes(x = Gender, reorder(Gender), y = AvgUnempRate, fill = Industry)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Average Unemployment Rate') +
  xlab('Gender') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  #scale_fill_manual(values=c( "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  labs(title = "Arts, Leisure and Hospitality experienced the highest unemployment for both genders", 
       subtitle = "More men worked in the Retail industry than women,, but unemployment rates were higher for women", 
       x = "Gender", y = "Avg Unemployment Rate (2020)", 
       tag = , caption = "") 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Show plots
plot3
plot4 
```

\

- 72.42% of women worked in `Retail` or `Arts, Leisure and Hospitality`   

- Compared to men, women experienced a higher spike in unemployment during the COVID-19 shutdown.    

- Unemployment rates remained higher for women even after the partial re-opening in May.  

- Women experienced higher unemployment across all industries, with the exception of `Manufacturing`. `Retail` was the second most affected industry, regardless of gender.      

- Retail employed 4.4% more men than women, but unemployment rates were higher for women.      

\

**By Education**

\

```{r, echo = FALSE, message=FALSE, warning=FALSE}

# Education by Gender

# Male Education
male_edu_2020 <- df_2020 %>% 
  filter(Gender == "Male",) %>%
  group_by(Education = as_factor(Education)) %>%
  summarize(n = sum(WTFINL)) %>%
  mutate(PercentMale = n/sum(n))
# male_edu_2020

# Female Education
female_edu_2020 <- df_2020 %>% 
  filter(Gender == "Female",) %>% 
  group_by(Education = as_factor(Education)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
# female_edu_2020

female_edu_2020$Education <- factor(female_edu_2020$Education, levels =c("GraduateDegreeOrHigher", "NoHighschool", "BachelorDegree", "SomeCollegeAssociateDegree","HighschoolNoCollege" ))

plot5 <- ggplot(data = female_edu_2020, aes(x = Education, y = PercentFemale, fill = Education)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('PercentFemale') +
  xlab('Education') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  labs(title = "Women's Education Level (2020)", 
       subtitle = "Over 70% of women have less than an Associate Degree ", 
       x = "Education", y = "Percent Women", 
       tag = , caption = "")

# Unemployemnt by Edu level
women_unemp_2020 <- df_2020 %>% 
  filter(Gender == "Female") %>% 
  group_by(MONTH, Education = as_factor(Education)) %>% 
  summarize(Female_AvgUnempRate = 1 - (weighted.mean(Employed, w = WTFINL)),
            Female_AvgEmpRate = weighted.mean(Employed, w = WTFINL))

plot6 <- ggplot(data = women_unemp_2020, aes(x = MONTH, y = Female_AvgUnempRate, color = Education)) +
  #geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  xlim('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec') +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values=c("#b3cde0", "#6497b1", "#005b96", "darkgrey", "#011f4b"))+
  theme_clean() +
   geom_vline(aes(xintercept = 4 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Apr", y = .03, label = 'COVID Shutdown', vjust= .5, hjust = 1, fill = 'white', size = 7/.pt, color = "gray29" ) +
  geom_vline(aes(xintercept = 7 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Jul", y = .03, label = 'Reopening Paused', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  geom_vline(aes(xintercept = 5 ), linetype = 'dashed', color = 'lightblue3') +
  annotate(geom = 'text', x = "May", y = .06, label = 'Partial Reopening', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  labs(title = "Women with lower education levels experienced a higher unemployment \nduring COVID-19 shutdown", 
       x = "Month", y = "Avg Unemployment Rate", 
       tag = , caption = "")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot5
plot6
```

\

- Over 70% of women have less than an Associate Degree  

- Women with lower education levels experienced a higher unemployment during COVID-19 shutdown   

- 26.60% of women have graduate or higher education.  They experienced a lower peak in unemployment during the COVID-19 shutdown.   

- Unemployment decline after the partial opening in May, but increased in November and December.    

\

### Summary of Exploratory Findings

\

We examined employment data for 4 different industries `Retail`, `Manufacturing`, `Information` and `Arts, Leisure, and Hospitality` to understand the scope of employment as it pertains to gender, age, race, and education.
Data from 2019 to 2020 showed:  

+ 7.5% average unemployment rate
+ Majority of Minority Americans worked in `Retail` or `Arts, Leisure and Hospitality`
+ 72% of Mixed-Race Americans, 68% of Latino Americans, and 67% of Black Americans
+ 72% of women worked in `Retail` or `Arts, Leisure and Hospitality` compared to 55%  of men
+ 43% of women have less than a high school education compared to 34% of men
+ Among women who are head of household 21% are between the age of 20 and 29

Looking at unemployment for 2020, we found that during the COVID-19 shutdown:  

+ Black Americans experienced the most significant increase in unemployment
+ Unemployment was the highest for men and women  who worked in  the `Arts, Leisure and Hospitality` industry
+ Women experienced higher unemployment than men across all industries
+ Retail industry employed 4.4% more men than women, but unemployment rates were higher for women
+ Over 70% of women have less than an Associate Degree
+ Women with lower education levels experienced a higher unemployment

\

## Models

\

Based on the findings in the exploratory analysis, we will examine the effect of `Gender` on the probability of being employed. Additionally, we will look at the effect of the following interactions on the probability of being employed:     

+ `Gender` and `Race`    
+ `Gender` and `Age`    
+ `Gender` and `Education`    

\

First, we run a model looking solely at the effect of  `Gender` on the probability of being employed.

\

$$ Employed_i= \beta_0 + \beta_1Gender_i $$
\

The second model measures the effect of the interaction between `Gender` and `Race` on the probability of being employed. 

\

$$ Employed_i= \beta_0 + \beta_1Gender_i\times Race_i$$

\

The third model measures the effect of the interaction between `Gender` and `Age` on the probability of being employed. 

\

$$ Employed_i= \beta_0 + \beta_1Gender_i\times Age_i$$

\

The last model measures the effect of the interaction between `Gender` and `Education`  on the probability of being employed. 

\

$$ Employed_i= \beta_0 + \beta_1Gender_i\times Education_i   $$

\

### Models Results

\

***Model 1:*** The effect of `Gender` on the probability of being employed  

$$ Employed_i= \beta_0 + \beta_1Gender_i $$

```{r, echo = FALSE, message=FALSE, warning=FALSE}

# Baseline model : Model 1: The effect of `Gender` on the probability of being employed

Model_1 <- lm_robust(Employed ~ Gender, data = df_2020, weights = WTFINL)

# Perform hypothesis test that all coefficients are equal to 0
# linearHypothesis(Model_1, matchCoefs(Model_1, "Gender"))

# Look at model
# export_summs(Model_1)
export_summs(Model_1,
             model.names = c("Gender"),
             coefs = c("Female" = "GenderFemale"),
             robust = TRUE)
```

\

+ Holding all else constant, women have a 2 percentage points less probability of being employed relative to men.  The estimate in this model is statistically significant at the 0.001 level.  

\

***Model 2:*** The effect of the interaction between `Gender` and `Race` on the probability of being employed.  

$$ Employed_i= \beta_0 + \beta_1Gender_i\times Race_i$$
```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Model 2: The effect of the interaction between `Gender` and `Race` on the probability of being employed. 

Model_2 <- lm_robust(Employed ~ Gender +  Race, data = df_2020, weights = WTFINL)
Model_2a <- lm_robust(Employed ~ Gender * Race, data = df_2020, weights = WTFINL)

# Perform hypothesis test that all coefficients are equal to 0
# linearHypothesis(Model_2, matchCoefs(Model_2, "Gender"))
# linearHypothesis(Model_2a, matchCoefs(Model_2a, "Gender"))
# # Look at model
export_summs(Model_2, Model_2a,
             model.names = c("Gender,Race", "Gender:Race"),
             coefs = c("Female" = "GenderFemale",
                       "Black" = "RaceBlack",
                       "Latino" = "RaceLatino",
                       "MixedRace" = "RaceMixedRace",
                       "White" = "RaceWhite",
                       "Female:Black" = "GenderFemale:RaceBlack",
                       "Female:Latino" = "GenderFemale:RaceLatino",
                       "Female:MixedRace" = "GenderFemale:RaceMixedRace",
                       "Female:White" = "GenderFemale:RaceWhite"),
             scale = TRUE, robust = TRUE)
```

\

+ Controlling for gender, Black Americans are 5 percentage points less likely to be employed compared to Asian Americans.  Whereas White Americans are 2% more likely to be employed relative to Asian Americans.    
  
+ Relative to Asian American men, Black American men are 6 percentage points less likely to be employed.  
In contrast, White American men are 2 percentage points more  likely to be employed than Asian American men.  
 
+ Black American women (-0.04 -0.06 +0.03 = -0.07) are 1 percentage point less likely to be employed compared to Black American men (-0.06).    

+ The estimates in these models are statistically significant at the 0.001 level. 

\

+ **Higher probability  of employment:**  White American men  
+ **Lower  probability  of employment:**  Black American women  

\

***Model 3:*** The effect of  the interaction between `Gender` and `Age` on the probability of being employed.  

$$ Employed_i= \beta_0 + \beta_1Gender_i\times Age_i$$

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Model 3: The effect of  the interaction between `Gender` and `Age` on the probability of being employed. 

Model_3 <- lm_robust(Employed ~ Gender + AgeGroup, data = df_2020, weights = WTFINL)
Model_3a <- lm_robust(Employed ~ Gender * AgeGroup, data = df_2020, weights = WTFINL)

# Perform hypothesis test that all coefficients are equal to 0
# linearHypothesis(Model_3, matchCoefs(Model_3, "Gender"))
# linearHypothesis(Model_3a, matchCoefs(Model_3a, "Gender"))

# Look at model
export_summs( Model_3, Model_3a,
             model.names = c("Gender,Age", "Gender:Age"),
             coefs = c("Female" = "GenderFemale",
                       "Age:30to39" = "AgeGroup30to39",
                       "Age:40to49" = "AgeGroup40to49",
                       "Age:50to59" = "AgeGroup50to59",
                       "Age:over60" = "AgeGroupover60",
                       "Age:under20" = "AgeGroupunder20",
                       "Female:Age 30to39" = "GenderFemale:AgeGroup30to39",
                       "Female:Age 40to49" = "GenderFemale:AgeGroup40to49",
                       "Female:Age 50to59" = "GenderFemale:AgeGroup50to59",
                       "Female:Age over60" = "GenderFemale:AgeGroupover60",
                       "Female:Age under20" = "GenderFemale:AgeGroupunder20"),
             scale = TRUE, robust = TRUE)
```

\

+ Holding gender constant, those who are 40 years old or older are 6 percentage points more likely to be employed compared to those who are 20 to 29 years old.    
+ Among those men and women between 20 to 29 years old, women are 1 percentage point less likely to be employed.   
+ Men between the age of 50 and 59 are 7 percentage points more likely to be employed compared to men between 20 to 29 years old.    

+ Women between the age of 50 and 59 are 4 percentage points more likely to be employed compared to women between 20 to 29 years old.    

+ The estimates in these models are statistically significant at the 0.001 level.    
\

+ **Higher probability  of employment:**  Men between the age of 50 and 59  
+ **Lower  probability  of employment:**  Women between 20 to 29 years old  
+ **Note:**  Among women who are head of household 21% are between the age of 20 and 29  

\

***Model 4:*** The effect of the interaction between `Gender` and `Education`  on the probability of being employed. 
$$ Employed_i= \beta_0 + \beta_1Gender_i\times Education_i   $$

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Model 4 : The effect of the interaction between `Gender` and `Education`  on the probability of being employed. 

Model_4 <- lm_robust(Employed ~ Gender + Education, data = df_2020, weights = WTFINL)
Model_4a <- lm_robust(Employed ~ Gender * Education, data = df_2020, weights = WTFINL)

# Perform hypothesis test that all coefficients are equal to 0
# linearHypothesis(Model_4, matchCoefs(Model_4, "Gender"))
# linearHypothesis(Model_4a, matchCoefs(Model_4a, "Gender"))

# # Look at model
export_summs(Model_4, Model_4a,
             model.names = c("Gender,Education", "Gender:Education"),
             coefs = c("Female" = "GenderFemale",
                       "GraduateDegreeOrHigher" = "EducationGraduateDegreeOrHigher",
                       "HighschoolNoCollege" = "EducationHighschoolNoCollege",
                       "NoHighschool" = "EducationNoHighschool",
                       "SomeCollegeAssociateDegree" = "EducationSomeCollegeAssociateDegree",
                       "Female:GraduateDegreeOrHigher" = "GenderFemale:EducationGraduateDegreeOrHigher",
                       "Female:HighschoolNoCollege" = "GenderFemale:EducationHighschoolNoCollege",
                       "Female:NoHighschool" = "GenderFemale:EducationNoHighschool",
                       "Female:SomeCollegeAssociateDegree" = "GenderFemale:EducationSomeCollegeAssociateDegree"),
             scale = TRUE, robust = TRUE)
```

\

+ Controlling for gender, those with no high school education has a 5 percentage points less probability of being employed relative to having a Bachelor's degree.       
+ Comparing men and women with Bachelor’s  degrees, women are 3 percentage points less likely to be employed.      

+ Men with NO high school education are 6 percentage points less likely to be employed compared to men with Bachelor’s degrees.    

+ Women with NO high school education ( -0.03 -0.06 + 0.02) are 7 percentage points less likely to be employed compared to men with Bachelor’s degrees.    

+ Women with graduate degrees  ( -0.03 +0.02 + 0.02= 0.01) are 7 percentage points MORE (0.01 – (-0.06)) likely to be employed compared to men with NO high school education (-0.06).   

+ Among men and women with graduate or higher educations, men are 1 percentage point more likely to be employed.    

+ Comparing women with NO high school education (-0.03 - 0.06 + 0.02 = -0.07)  and women with Associates degrees 
(-0.03 -0.03 +0.01 = -0.05 ), those with Associates degrees are 2 percentage point more likely to be employed.   

+ The estimates in these models are statistically significant at the 0.001 level.   
\

+ **Higher probability of employment:**  Men with Bachelor’s or higher degrees   
+ **Lower  probability of employment:**  Women with lower than Associates degrees 

\
  
## Summary of Findings

\

Who has money to spend?  

We examined the 2019 to 2020 Current Population Survey data for 4 different industries to understand the scope of employment as it pertains to gender, age, race, and education.  Within the `Retail`,  `Manufacturing`, `Information` and `Arts, Leisure, and Hospitality` industries, we found that on average over 72% of women and over 67% of Minority Americans worked in `Retail` or `Arts, Leisure and Hospitality`.   Overall, unemployment was higher among women and Black Americans.  In addition, the `Arts, Leisure and Hospitality` industry suffered the most unemployment during the COVID-19 shutdown.       

Our assumption is that those who are working during the COVID-19 pandemic have more spending power than those who are unable to work.  And individuals with a higher probability of being employed are more likely to work during the COVID-19 shutdown.    

Our analysis showed that individuals with highest probability of employment are:  

+ White American men    
+ Men between the age of 50 and 59    
+ Those with a graduate or higher degree    
   
Individuals with the lowest probability of being employed are:      

+ Black American women  
+ Women between the age of 20 and 29  
+ Women who have lower than Associates degrees  

Overall, Black American women, women with no college education, or women between the age of 20 and 29 have lower  spending power.  We also observed that among women who are head of household, 21% are between the age of 20 and 29.   Given that on average 72% of women worked in  `Retail` or `Arts, Leisure and Hospitality` they experienced the most unemployment.     
 
\
