---
title: "EDA: Demographics"
author: "Elizabeth Tran"
date: "3/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# theme_clean <- function(x) {
#   #theme_minimal() + 
#   theme(panel.border = element_blank(),
#         panel.background = element_blank(),
#         axis.line = element_line(color = "gray"),
#         axis.ticks.y = element_blank(),
#         axis.ticks.x = element_blank(),
#         plot.title = element_text(color = "gray29", size = 11, face = "plain"),
#         plot.subtitle = element_text(color = "gray29", size = 9, face = "plain"),
#         axis.text.x = element_text(color = "#595959", size = 6.5),
#         axis.text.y = element_text(color = "#595959", size = 8),
#         axis.title = element_text(color = "#595959", size = 10, face = "plain"),
#         legend.title = element_text(color = "#595959", size = 10, face = "plain"),
#         legend.text = element_text(color = "#595959", size = 10, face = "plain"),
#         #panel.background = element_blank()
#         )
# }
```



## Libraries

```{r}
library(tidyverse)
library(jtools)
library(vtable)
library(car)
library(estimatr)
library(lubridate)
library(readr)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(ipumsr) # IPUMS data mapping
# vignette("value-labels", package = "ipumsr")
library(gridExtra)
library(fastDummies)

```

## NOTE: 
To load data, you must download both the extract's data and the DDI and also set the working directory to the folder with these files (or change the path below).  

If (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

ddi <- read_ipums_ddi("cps_00006.xml")
data <- read_ipums_micro(ddi)



## IMPORT DATA
```{r}


ddi <- read_ipums_ddi("cps_00006.xml")
data <- read_ipums_micro(ddi)

StateCode <- read_csv('State_FIPS.csv')

StateCode <- StateCode %>% 
  select(State, STATEFIP)

Industries <- read_csv('indnames.csv')
# unique(Industries$indname)

Occupations <- read_csv('Occupation.csv')

Industries <- Industries %>%
  mutate(Industry = indname, IND = ind) %>%
  select(-indname, -ind) %>% 
  filter(Industry %in% c('Manufacturing', 
                         'Retail Trade', 
                         'Information',
                         'Arts, Entertainment, and Recreation, and Accommodation and Food Services')) %>% 
  mutate(Industry = case_when(Industry == 'Retail Trade' ~ 'Retail',
                              Industry == 'Manufacturing' ~ 'Manufacturing',
                              Industry == 'Information' ~ 'Information',
                              Industry == 'Arts, Entertainment, and Recreation, and Accommodation and Food Services' ~ 'Arts, Leisure and Hospitality'))
  

# vtable(data, lush = TRUE)

```





## DATA CLEANING
```{r}
class_converted <- data %>%
  mutate(ASECFlag = as_factor(lbl_clean(ASECFLAG)),
         FoodStampRecipient = as_factor(lbl_clean(FOODSTMP)),
         Sex = as_factor(lbl_clean(SEX)),
         LaborForce = as_factor(lbl_clean(LABFORCE)),
         COVIDTelework = as_factor(lbl_clean(COVIDTELEW)),###### has NIU
         COVIDUnableToWork = as_factor(lbl_clean(COVIDUNAW)),###### has NIU
         COVIDUnableToLookForWork = as_factor(lbl_clean(COVIDLOOK)),###### has NIU
         FamilyIncome = as_factor(lbl_clean(FAMINC))
         ) %>% 
  select(-c(ASECFLAG, FOODSTMP, SEX, COVIDTELEW, COVIDUNAW, COVIDLOOK, FAMINC, LABFORCE))

df <- class_converted %>% 
  select(-SERIAL) %>% 
  mutate(AgeGroup = case_when(AGE < 20 ~ 'under20',
                               AGE >= 20 & AGE < 30 ~ '20to29',
                               AGE >= 30 & AGE < 40 ~ '30to39',
                               AGE >= 40 & AGE < 50 ~ '40to49',
                               AGE >= 50 & AGE < 60 ~ '50to59',
                               AGE >= 60 ~ 'over60')) %>% 
  mutate(Regions = case_when(REGION == 11 | REGION == 12 ~ 'Northeast',
                             REGION == 21 | REGION == 22 ~ 'Midwest',
                             REGION == 31 | REGION == 32 | REGION == 33 ~ 'South',
                             REGION == 41 | REGION == 42 ~ 'West',
                             REGION == 97 ~ 'Unknown')) %>% 
  select(-REGION) %>% 
  mutate(MetroArea = case_when(METRO %in% c(0, 4, 9) ~ 'Others',
                               METRO == 1 ~ 'Rural',
                               METRO == 2 ~ 'Urban',
                               METRO == 3 ~ 'Suburb')) %>% 
  select(-METRO) %>% 
  mutate(Race = case_when(HISPAN == 0 & RACE == 100 ~ 'White',
                          HISPAN == 0 & RACE == 200 ~ 'Black',
                          HISPAN == 0 & RACE %in% c(650, 651, 652) ~ 'Asian',
                          HISPAN == 0 & !(RACE %in% c(100, 200, 650, 651, 652)) ~ 'MixedRace',
                          HISPAN != 0 ~ 'Latino')) %>% 
  select(-ASIAN, -HISPAN, -RACE) %>%
  mutate(Married = ifelse(MARST == 1, 1, 0)) %>% 
  select(-MARST) %>% 
  left_join(StateCode, by = 'STATEFIP') %>% 
  select(-STATEFIP) %>% 
  inner_join(Industries, by = 'IND') %>% 
  select(-IND) %>% 
  mutate(Education = case_when(EDUC >= 10 & EDUC <= 72 ~ 'NoHighschool',
                               EDUC == 73 ~ 'HighschoolNoCollege',
                               EDUC %in% c(80, 81, 90, 100, 110, 120, 121, 122, 91, 92) ~ 'SomeCollegeAssociateDegree',
                               EDUC == 111 ~ 'BachelorDegree',
                               EDUC %in% c(123,124,125) ~ 'GraduateDegreeOrHigher',
                               EDUC %in% c(0, 1, 2, 999) ~ 'NA')) %>% 
  
  select(-EDUC) %>% 
  filter(Education != 'NA') %>% 
  mutate(Employed = case_when(EMPSTAT <= 1 | EMPSTAT >= 30 ~ 'Others', 
                               EMPSTAT >= 10 &  EMPSTAT < 20 ~ '1',
                               EMPSTAT >= 20 &  EMPSTAT < 30 ~ '0'
                               )) %>% 
  select(-EMPSTAT) %>%
  mutate(WorkStatus = case_when(WKSTAT %in% c(10,11,14,15) ~ 'Full Time',
                                 WKSTAT %in% c(12,20,21,22,40,41) ~ 'Part Time',
                                 WKSTAT %in% c(13,42,50,60) ~ 'Unemployed',
                                 WKSTAT == 99 ~ 'NA' )) %>% 
  select(-WKSTAT) %>% 
  mutate(UnemploymentReason = ifelse(WHYUNEMP==1, 'Layoff', 'Other')) %>% 
  select(-WHYUNEMP) %>% 
  mutate(PartTimeReason = case_when(!(WHYPTLWK %in% c(52, 60, 101, 121, 122)) ~ 'Others',
                                    WHYPTLWK == 52 ~ 'Job Terminated',
                                    WHYPTLWK == 60 ~ 'Could Only Find PT',
                                    WHYPTLWK == 101 ~ 'Health Limitation',
                                    WHYPTLWK == 121 ~ 'Childcare Problem',
                                    WHYPTLWK == 122 ~ 'Other Personal Obligations')) %>% 
  select(-WHYPTLWK) %>% 
  left_join(Occupations, by = 'OCC') %>% 
  select(-OCC) %>% 
  filter(Employed != 'Others') %>% 
  filter(!is.na(WTFINL))

df$Employed <- as.numeric(df$Employed)

df <- dummy_cols(df, select_columns = 'Race')
  

vtable(df, lush=TRUE)

```

<br>
<br>


## OVERVIEW OF DEMOGRAPICS (2019 to 2020)  

### EMPLOYMENT  
```{r}
demographics <-  df

# Employment vs unemployment rate
demographics <-  df %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics

##  Proportion of the industries
demographics <-  df %>% 
  group_by(Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics

##  What does employment look like for each industry?  
demographics <- df %>% 
  group_by(Industries = as_factor(Industry)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))

demographics

ggplot(data = demographics, aes(x = Industries, y = AvgUnemploymentRate, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  coord_flip() +
  ggtitle("Average Unemployment Within Each Industry")+
  theme_bw()
  
  

```

+ ***Within the 4 industries, average  unemployment rate is 7.5% on from 2019 to 2020.***  
+ ***Proportion of the industries: ***  
  + 33.31%  ***Retail*** ,   (within Retail avg unemployment rate is 6.64%)       
  + 31.61%  Manufacturing         
  + 29.45%  Arts, Leisure and Hospitality,  (within ALH avg unemployment rate is 11.8% ***highest***)  
  + 5.62 %  Information,   (within Information avg unemployment rate is 5.49% ***lowest***)        

<br>
<br> 



### AGE  
```{r}

# AgeGroup Within the industries
demographics <- df %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics

ggplot(data = demographics, aes(x = AgeGroup, y = percent, fill = AgeGroup)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  coord_flip() +
  ggtitle("Age Group within the 4 industries")+
  theme_bw()


# Employment Within the AgeGroup
demographics <- df %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))

demographics

ggplot(data = demographics, aes(x = AgeGroup, y = AvgUnemploymentRate, fill = AgeGroup)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  coord_flip() +
  ggtitle("Unemployment Rate within  Age Group")+
  theme_bw()


# Where are they working?
demographics <- df %>% 
  group_by(AgeGroup = as_factor(AgeGroup),
           Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
demographics  

ggplot(data = demographics, aes(x = AgeGroup, y = Percent, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Age Group') +
  ggtitle("Where Are They working? (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))


# What does uemployment look like by industry?

demographics <- df %>%
  group_by(AgeGroup = as_factor(AgeGroup),
           Industries = as_factor(Industry)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics

ggplot(data = demographics, aes(x = AgeGroup, y = AvgUnemploymentRate, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  xlab('AgeGroup') +
  ggtitle("Unemployment by Age Within Industry (2019-2020)")+  
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))


```

+ ***Within all 4 industries, observations consist of:***  
+ 26.29% are  20to29 (largest group).  This group has average unemployment rate of  10.30%.         
+ 7.89%  are  under20 (smallest group).   
  + ***This age group slighly higher average unemployment rate of 10.50%.***    
  + ***59.44% of those under20 worked in 	Arts, Leisure and Hospitality***   
+ Those who are 40 and older have the highest level of employment       


<br>
<br> 

### GENDER  
```{r}
# Gender Within the industries
demographics <- df %>% 
  group_by(Gender = as_factor(Sex)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics

# Unmployment:  
demographics <- df %>% 
  group_by(Gender = as_factor(Sex)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics


# Which indsutry has the highest female proportion?
demographics <- df %>% 
  filter(Sex=="Female") %>% 
  group_by(Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n))
demographics

# what does Female  & unemployment proportion look like in each industry? 
demographics <- df %>% 
  group_by(Industries = as_factor(Industry),
           Gender = as_factor(Sex)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics

demographics <- df %>% 
  group_by(Industries = as_factor(Industry),
           Gender = as_factor(Sex)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics


ggplot(data = demographics, aes(x = Gender, y = AvgUnemploymentRate, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  xlab('Gender') +
  ggtitle("Unemployment by Gender Within Industry (2019-2020)")+  
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))

```
***Gender within the industries:  ***     

***Male***    

+ made up 57.43% of the observations and had 6.96% avg unemployment        
+ They work mostly in Manufacturing and Information    
+ Manufacturing employed 70.40%  Male  (***highest***)      

***Female***      

+ made up  42.57% of observations and had 8.34% avg unemployment        
+ They work mostly in Retail and and Arts, Leisure and Hospitality industries (72.70% combined)     
+ These industries has the highest unemployment      
+ Arts, Leisure and Hospitality employed 51.01% women (***highest industry for women***)    
+ Information employed 5.330%  only females (***lowest industry for women***)    

***In Retail: ***    

+ There are 4.42% more men working in the Retail industry than women,
but unemployment is higher for women (7.22%)        



<br>
<br> 


### ETHNICITY  
```{r}
# Ethnicity Within the Industries
demographics <- df %>% 
  group_by(Race = as_factor(Race)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics

demographics <- df %>% 
  group_by(Race = as_factor(Race)) %>% 
  summarize(AvgUnemploymentRate = 1 - (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics


# Where are they working?
demographics <- df %>% 
  group_by(Race = as_factor(Race),
           Industries = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
demographics


ggplot(data = demographics, aes(x = Race, y = Percent, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Race') +
  ggtitle("Where Are They working? (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))


# What does uemployment look like by industry?

demographics <- df %>%
  group_by(Race = as_factor(Race),
           Industries = as_factor(Industry)) %>% 
  summarize(AvgUnemploymentRate = 1- (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics

ggplot(data = demographics, aes(x = Race, y = AvgUnemploymentRate, fill = Industries)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  xlab('Race') +
  ggtitle("Unemployment by Age Within Industry (2019-2020)")+  
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))
```

***Ethnicity made up of: ***    

+ 59.17% White & had the ***lowest avg unemployment*** (6.28%)      
+ 19.73% Latino & had 8.41% avg unemployment      
+ 11.69% Black & had the ***highest avg unemployment*** (12.15%)    
+  
+ 67.10% of Black worked in Retail and ALH industries and had a total avg unemployment rate of 28.2%  
+ 68.40% of Latino worked in Retail and ALH industries and had a total avg unemployment rate of 18.76%  
+ 72.32% of MixedRace worked in Retail and ALH industries



<br>
<br> 

### EDUCATION 
```{r}
# Education Within the Industries
demographics <- df %>% 
  group_by(Education = as_factor(Education)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(percent = n/sum(n))
demographics



# Where are they working?
demographics <- df %>% 
  group_by(Education = as_factor(Education),
           Industry = as_factor(Industry)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
demographics

ggplot(data = demographics, aes(x = Education, y = Percent, fill = Industry)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Education Level') +
  ggtitle("Where Are They working? (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))



# Unemployemnt by Edu level
demographics <- df %>% 
  group_by(Education = as_factor(Education)) %>% 
  summarize(AvgUnemploymentRate = 1 - (weighted.mean(Employed, w = WTFINL)),
            AvgEmpRate = weighted.mean(Employed, w = WTFINL))
demographics

ggplot(data = demographics, aes(x = Education, y = AvgUnemploymentRate, fill = Education)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('AvgUnemploymentRate') +
  xlab('Education Level') +
  ggtitle("Unemployment (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))



# Education by Age 
demographics <- df %>% 
  group_by(Education = as_factor(Education),
           AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
demographics

ggplot(data = demographics, aes(x = AgeGroup, y = Percent, fill = Education)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Education Level') +
  ggtitle("Education by Age (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))

# Education by Gender

demographics <- df %>% 
  group_by(Education = as_factor(Education),
           Gender = as_factor(Sex)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(Percent = n/sum(n))
demographics

ggplot(data = demographics, aes(x = Gender, y = Percent, fill = Education)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Education Level') +
  ggtitle("Education by Gender (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0))

# Women's Education Level
demographics <- df %>% 
  filter(Sex == "Female") %>% 
  group_by(Education = as_factor(Education)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
demographics

```


***Education: ***  

+ 32.94% had only HighschoolNoCollege
  + The worked mostly in Mfgr, *Retail (35.76%)*  &  ALH      
  +  (8.53% AvgUnemploymentRate)   
  
+ 29.94% had SomeCollegeAssociateDegree   
  + The worked mostly in Mfgr, ***Retail (37.3%)***  &  ALH   

+ 11.15%  had NoHighschool 
  + 47.40% worked in Arts, Leisure and Hospitality   
+ ( 9.80% AvgUnemploymentRate *highest*)  

**74.20% of women have less than a SomeCollegeAssociateDegree**





### WOMEN
```{r}
demographics <- df %>% 
  filter(Sex=="Female")
# sumtable(employment_status_gender5)

# # Which indsutry has the highest female proportion?
demographics <- df %>%
  filter(Sex=="Female") %>%
  group_by(Industries = as_factor(Industry)) %>%
  summarize(n = sum(WTFINL)) %>%
  mutate(PercentFemale = n/sum(n))
demographics


# # Women's Education Level
demographics <- df %>%
  filter(Sex == "Female") %>%
  group_by(Education = as_factor(Education)) %>%
  summarize(n = sum(WTFINL)) %>%
  mutate(PercentFemale = n/sum(n))
demographics

#Women's Marital Status
demographics <- df %>% 
  filter(Sex == "Female") %>% 
  group_by(Married = as_factor(Married)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
demographics

# Women as Head of Household
demographics <- df %>% 
  filter(PERNUM == 1) %>% #head of household
  group_by(Gender = as_factor(Sex)) %>% 
  summarize(n = sum(HWTFINL)) %>% 
  mutate(Percent = n/sum(n)) 
demographics

# Women's ageGroup
demographics <- df %>% 
  filter(Sex == "Female") %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
demographics

# Women's ageGroup & Head of household
demographics <- df %>% 
  filter(PERNUM == 1,
         Sex == "Female") %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(HWTFINL)) %>% 
  mutate(PercentFemaleHeadHousehold = n/sum(n)) 
demographics

# Women's Ethnicity
demographics <- df %>% 
  filter(Sex == "Female") %>% 
  group_by(Race = as_factor(Race)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(PercentFemale = n/sum(n)) 
demographics


```
***WOMEN***

+ 72.70%  of women worked in Retail and ALH industries      
+ 74.20% % of women have less than a SomeCollegeAssociateDegree    
+ 61.16% of women are single      
+ 39.81% of population are Female head of household ( % spread evenly over ages 20-59)    
+ 27.26% of women are 20to29    



<br>
<br> 


### WOMEN AFTER COVID-19  

```{r}
postCOVID <- df %>% 
filter(COVIDUnableToWork == "Yes") %>% 
  group_by(Sex = as_factor(Sex)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDUnableToWorkPercent  = n/sum(n)) 
postCOVID

postCOVID <- df %>% 
  filter(COVIDUnableToWork == "Yes",
         Sex=="Female") %>% 
  group_by(Race = as_factor(Race)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDUnableToWorkPercentFemale = n/sum(n)) 
postCOVID

postCOVID <- df %>% 
  filter(COVIDUnableToWork == "Yes",
         Sex=="Female") %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDUnableToWorkPercentFemale = n/sum(n)) 
postCOVID

postCOVID <- df %>% 
  filter(COVIDUnableToWork == "Yes",
         Sex=="Female") %>% 
  group_by(Education = as_factor(Education)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDUnableToWorkPercentFemale = n/sum(n)) 
postCOVID

```
 ***WOMEN AFTER COVID-19***
 
 + Of 18054 COVIDUnableToWork observations, 47.18% are women    
 + Of women who are COVIDUnableToWork, 30.73% are in the age group 20to29      
 + Of women who are COVIDUnableToWork,  63% of women have less than  SomeCollegeAssociateDegree   

<br>
<br> 



### COVID WHO CAN WORK FROM HOME?
```{r}

postCOVID <- df %>% 
  filter(COVIDTelework == "Yes") %>% 
  group_by(Gender = as_factor(Sex)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDTeleworkPercent = n/sum(n)) 
postCOVID

postCOVID <- df %>% 
  filter(COVIDTelework == "Yes") %>% 
  group_by(AgeGroup = as_factor(AgeGroup)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDTeleworkPercent = n/sum(n)) 
postCOVID

postCOVID <- df %>% 
  filter(COVIDTelework == "Yes") %>% 
  group_by(Race = as_factor(Race)) %>% 
  summarize(n = sum(WTFINL)) %>% 
  mutate(COVIDTeleworkPercent = n/sum(n)) 
postCOVID

```

***COVID WHO CAN WORK FROM HOME?***   

+ Of 18054 COVIDTelework observations (those who can work from home), 56.55% men and 43.4% are women.  
+ Only 1.02% of people under20 can work from home.  
+ 




<br>
<br> 
 

### COVID UNABLE TO LOOK FOR FOR 
***Skip:  Variable has only NA and NIU***
```{r}
# postCOVID <- df %>%
#   group_by(COVIDUnableToLookForWork = as_factor(COVIDUnableToLookForWork)) %>%
#   summarize(n = sum(WTFINL)) %>%
#   mutate(COVIDUnableToLookForWorkPercent = n/sum(n))
# postCOVID
```

<br>
<br> 


### UNEMPLOYMENT OVER 2 YEARS
```{r}

demographics <- df %>% 
  mutate(mo = as.numeric(MONTH), 
         Yr = as.numeric(YEAR),
         YearMonth = Yr*100 + mo) %>%   #create a year-month variable 
  group_by( YearMonth, Race ) %>% 
  summarize(AvgUnemploymentRate = 1 - (weighted.mean(Employed, w = WTFINL)))
            
demographics$YearMonth <- as.factor(demographics$YearMonth)

ggplot(data = demographics, aes( x = YearMonth, y = AvgUnemploymentRate,  color = Race, shape = Race)) +
  geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Percent') +
  xlab('Education Level') +
  ggtitle("Unemployment by Race (2019-2020)")+
  theme(axis.text.x = element_text(angle = 90)) +

  theme(axis.text.x = element_text(angle = 0))

```




<br>
<br> 
### Modeling variables :  
race, education, gender, telework, unabletowork, Industries, region, metro   

Q1 Model: Region & Metro    
lm(Employed ~ retail + metro)  
lm(Employed ~ retail + region)  

Q2 Model: industry, interaction between race & gender  
lm(Employed ~ industry + COVIDUnableToWork)  
lm(employed ~ industry + COVIDTelework)  

Q3 Model:    
lm(Employed ~ Gender + eduation)  
lm(Employed ~ Gender + race)  


<br>
<br> 
<br>
<br> 

















