---
title: "OMSBA 5300 Data Translation Project"
author: "Elizabeth, John, Julius, Raymond, Schyuler"
date: "2/22/2021"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo = FALSE, message=FALSE, warning=FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```


## Libraries and set themes

```{r, message= FALSE, warning=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)
library(jtools)
library(vtable)
library(car)
library(estimatr)
library(lubridate)
library(readr)
library(ggplot2)
library(ggpubr)
library(fastDummies)
library(viridis)
library(gridExtra)
library(cowplot)

# Set themes




theme_clean <- function(x) {
  #theme_minimal() + 
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "gray"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(color = "gray29", size = 11, face = "plain"),
        plot.subtitle = element_text(color = "gray29", size = 9, face = "plain"),
        axis.text.x = element_text(color = "#595959", size = 6.5),
        axis.text.y = element_text(color = "#595959", size = 8),
        axis.title = element_text(color = "#595959", size = 10, face = "plain"),
        legend.title = element_text(color = "#595959", size = 10, face = "plain"),
        legend.text = element_text(color = "#595959", size = 10, face = "plain"),
        strip.background = element_rect(colour="White", fill="White")
        )
}
```

## Import Data

```{r, message = FALSE, warning=FALSE, echo=FALSE}
# Set default working directory
setwd("~/MSBA/11_Econometrics/Data Translation Project")

if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

ddi <- read_ipums_ddi("cps_00006.xml")
data <- read_ipums_micro(ddi)

StateCode <- read_csv('State_FIPS.csv')

StateCode <- StateCode %>% 
  select(State, STATEFIP)

Industries <- read_csv('indnames.csv')

Industries <- Industries %>%
  mutate(Industry = indname, IND = ind) %>%
  select(-indname, -ind) %>% 
  filter(Industry %in% c('Manufacturing', 
                         'Retail Trade', 
                         'Information',
                         'Arts, Entertainment, and Recreation, and Accommodation and Food Services')) %>% 
  mutate(Industry = case_when(Industry == 'Retail Trade' ~ 'Retail',
                              Industry == 'Manufacturing' ~ 'Manufacturing',
                              Industry == 'Information' ~ 'Information',
                              Industry == 'Arts, Entertainment, and Recreation, and Accommodation and Food Services' ~ 'Arts, Leisure and Hospitality'))
  

# vtable(data)

```


## Data Cleaning

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
class_converted <- data %>%
  mutate(ASECFlag = as_factor(lbl_clean(ASECFLAG)),
         FoodStampRecipient = as_factor(lbl_clean(FOODSTMP)),
         Sex = as_factor(lbl_clean(SEX)),
         LaborForce = as_factor(lbl_clean(LABFORCE)),
         COVIDTelework = as_factor(lbl_clean(COVIDTELEW)),###### has NIU
         COVIDUnableToWork = as_factor(lbl_clean(COVIDUNAW)),###### has NIU
         COVIDUnableToLookForWork = as_factor(lbl_clean(COVIDLOOK)),###### has NIU
         FamilyIncome = as_factor(lbl_clean(FAMINC))
         ) %>% 
  select(-c(ASECFLAG, FOODSTMP, SEX, COVIDTELEW, COVIDUNAW, COVIDLOOK, FAMINC, LABFORCE))

df <- class_converted %>% 
  select(-SERIAL) %>% 
  mutate(AgeGroup = case_when(AGE < 20 ~ 'under20',
                               AGE >= 20 & AGE < 30 ~ '20to29',
                               AGE >= 30 & AGE < 40 ~ '30to39',
                               AGE >= 40 & AGE < 50 ~ '40to49',
                               AGE >= 50 & AGE < 60 ~ '50to59',
                               AGE >= 60 ~ 'over60')) %>% 
  mutate(Regions = case_when(REGION == 11 | REGION == 12 ~ 'Northeast',
                             REGION == 21 | REGION == 22 ~ 'Midwest',
                             REGION == 31 | REGION == 32 | REGION == 33 ~ 'South',
                             REGION == 41 | REGION == 42 ~ 'West',
                             REGION == 97 ~ 'Unknown')) %>% 
  select(-REGION) %>% 
  mutate(MetroArea = case_when(METRO %in% c(0, 4, 9) ~ 'Others',
                               METRO == 1 ~ 'Rural',
                               METRO == 2 ~ 'Urban',
                               METRO == 3 ~ 'Suburb')) %>% 
  select(-METRO) %>% 
  mutate(Race = case_when(HISPAN == 0 & RACE == 100 ~ 'White',
                          HISPAN == 0 & RACE == 200 ~ 'Black',
                          HISPAN == 0 & RACE %in% c(650, 651, 652) ~ 'Asian',
                          HISPAN == 0 & !(RACE %in% c(100, 200, 650, 651, 652)) ~ 'MixedRace',
                          HISPAN != 0 ~ 'Latino')) %>% 
  select(-ASIAN, -HISPAN, -RACE) %>%
  mutate(Married = ifelse(MARST == 1, 1, 0)) %>% 
  select(-MARST) %>% 
  left_join(StateCode, by = 'STATEFIP') %>% 
  select(-STATEFIP) %>% 
  inner_join(Industries, by = 'IND') %>% 
  select(-IND) %>% 
  mutate(Education = case_when(EDUC >= 10 & EDUC <= 72 ~ 'NoHighschool',
                               EDUC == 73 ~ 'HighschoolNoCollege',
                               EDUC %in% c(80, 81, 90, 100, 110, 120, 121, 122, 91, 92) ~ 'SomeCollegeAssociateDegree',
                               EDUC == 111 ~ 'BachelorDegree',
                               EDUC %in% c(123,124,125) ~ 'GraduateDegreeOrHigher',
                               EDUC %in% c(0, 1, 2, 999) ~ 'NA')) %>%  
  select(-EDUC) %>% 
  filter(Education != 'NA') %>% 
  mutate(Employed = case_when(EMPSTAT <= 1 | EMPSTAT >= 30 ~ 'Others', 
                               EMPSTAT >= 10 &  EMPSTAT < 20 ~ '1',
                               EMPSTAT >= 20 &  EMPSTAT < 30 ~ '0'
                               )) %>% 
  select(-EMPSTAT) %>%
  mutate(WorkStatus = case_when(WKSTAT %in% c(10,11,14,15) ~ 'Full Time',
                                 WKSTAT %in% c(12,20,21,22,40,41) ~ 'Part Time',
                                 WKSTAT %in% c(13,42,50,60) ~ 'Unemployed',
                                 WKSTAT == 99 ~ 'NA' )) %>% 
  select(-WKSTAT) %>% 
  mutate(UnemploymentReason = ifelse(WHYUNEMP==1, 'Layoff', 'Other')) %>% 
  select(-WHYUNEMP) %>% 
  mutate(PartTimeReason = case_when(!(WHYPTLWK %in% c(52, 60, 101, 121, 122)) ~ 'Others',
                                    WHYPTLWK == 52 ~ 'Job Terminated',
                                    WHYPTLWK == 60 ~ 'Could Only Find PT',
                                    WHYPTLWK == 101 ~ 'Health Limitation',
                                    WHYPTLWK == 121 ~ 'Childcare Problem',
                                    WHYPTLWK == 122 ~ 'Other Personal Obligations')) %>% 
  select(-WHYPTLWK, -OCC) %>% 
  filter(Employed != 'Others') %>% 
  filter(!is.na(WTFINL)) #check with Schyuler

df$Employed <- as.numeric(df$Employed)
  
vtable(df)

```


\


## How has retail fared relative to other industries?

\

### Data Exploration

\

#### Overall employment figures by industry, year over year

\


```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Explore number of people employed by industry
jobs_industry <- df %>% 
  group_by(Industry, YEAR, Employed) %>% 
  count()

# Convert to factor
jobs_industry$YEAR <- as.numeric(jobs_industry$YEAR)
jobs_industry$Industry <- as.factor(jobs_industry$Industry)
jobs_industry$Employed <- factor(jobs_industry$Employed)

jobs_industry$Employed <- factor(jobs_industry$Employed, levels = c("1", "0"), 
                  labels = c("Yes", "No"))

# Plot jobs by industry
p1 <- ggplot(data = jobs_industry)+
  aes(x = YEAR)+
  aes(y = n) +
  aes(group = Employed)+
  facet_wrap(~ Industry)+
  geom_line(aes(linetype = Employed, color = Employed)) +
  geom_point(aes(color = Employed)) +
  scale_x_continuous(breaks = c(2019, 2020))+
  scale_y_continuous(labels = scales::comma ) +
  scale_color_manual(values=c("#6497b1", "#011f4b")) +
  theme_clean() +
  theme(legend.key=element_blank())+
  labs(title = "Retail and Manufacturing are similarly represented in the data and show similar overall \njob number trends", 
       subtitle = "Arts, Leisure and Hosptility lost the most  total jobs due to COVID19", 
       x = "Year", y = "Total Employed", 
       tag = , caption = "Source: IPUMS CPS Monthly Data")

p1


  
```

\

- Overall job numbers show a decrease from 2019 to 2020 for all industries. 

- `Manufacturing` and `Retail` are similarly represented in the data and show similar trends. `Arts, Leisure and Hospitality` shows the most dramatic drop in employment numbers

- Albeit less represented in the total sample, the `Information` industry appears to be the less impacted in terms of overall jobs, ear over year

\


#### Unemployment figures by industry


\



```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Filter years
df_2019 <- df %>% 
  filter(YEAR == 2019)

df_2020 <- df %>% 
  filter(YEAR == 2020) 


df_2020_COVID <- df %>% 
  filter(YEAR == 2020 & (!MONTH %in% c(1, 2, 3)))


## Unemployment Rate by Industries
unemp_ind_2020 <- df_2020 %>% 
  group_by(Industry, MONTH, Sex, Race, Education) %>% 
  summarize(AvgUnempRate = 1 - weighted.mean(Employed, w = WTFINL)) %>% 
  mutate(Diff = AvgUnempRate - lag(AvgUnempRate))

p2 <- ggplot(data = unemp_ind_2020, aes(x = MONTH, y = AvgUnempRate, color = Industry, shape = Industry)) +
  #geom_point() +
  geom_smooth(method = 'loess', se = FALSE) +
  xlim('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec') +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values=c("darkgray", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme_clean() +
 geom_vline(aes(xintercept = 4 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Apr", y = .03, label = 'COVID Shutdown', vjust= .5, hjust = 1, fill = 'white', size = 7/.pt, color = "gray29" ) +
  geom_vline(aes(xintercept = 7 ), linetype = 'dashed', color = 'lightgrey') +
  annotate(geom = 'text', x = "Jul", y = .03, label = 'Reopening Paused', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  geom_vline(aes(xintercept = 5 ), linetype = 'dashed', color = 'lightblue3') +
  annotate(geom = 'text', x = "May", y = .06, label = 'Partial Reopening', vjust= .5, hjust = -.03, fill = 'white', size = 7/.pt, color = "gray29") +
  labs(title = "Unemployment rates rose dramatically at the onset of the COVID-19 Pandemic", 
       subtitle = "Pausing the reopening of economy kept retail from improving more quickly", 
       x = "Month", y = "Avg Unemployment Rate", 
       tag = , caption = "Source: IPUMS CPS Monthly Data")

p2
         
```

\


- `Arts, Leisure and Hospitality` is the only industry more affected than `Retail` in terms of unemployment rates. 

- The partial reopening helped all industries recover from their April unemployment highs, with the pause on the reopening stalling each industry's recovery. When compared to `Retail`, `Manufacturing` has a lower unemployment peak and less of a sustained, higher unemployment rate 

- The `Information` industry drop in unemployment rate really took off in July, outpacing `Retail`recovery by November. 

- In spite of the traditional peak season for the `Retail` industry, unemployment trend does not significantly improve in terms of unemployment rate. 

\


#### Unable to work due to COVID19

\


```{r, echo = FALSE, message=FALSE}
df_2020_UnableToWork <- df_2020 %>% 
  filter(COVIDUnableToWork != 'NIU') %>% 
  mutate(UnableToWorkCOVID = ifelse(COVIDUnableToWork == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)

## COVIDUnableToWork Breakdown by Employed
UnableToWorkInd_2020 <- df_2020_UnableToWork %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgUnableToWork = 1 - weighted.mean(UnableToWorkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))
UnableToWorkInd_2020$Employed <- factor(UnableToWorkInd_2020$Employed)


# Factor industry
UnableToWorkInd_2020$Industry <- factor(UnableToWorkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))

# Plot 

p9 <- ggplot(data = UnableToWorkInd_2020, aes(x = Employed, y = AvgUnableToWork, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgUnableToWork), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  ylab('Average Rate of Reported Not Able To Work Due to COVID') +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  facet_wrap(~ Industry)+
  theme_clean()+
  theme(legend.position="none") +
  labs(title = "Manufacturing reported the highest percentage of being unable to work due to COVID-19, \nfollowed by Retail ", 
       subtitle = "", 
       x = "Month", y = "% Reporting Unable to Work due to COVID-19", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 





df_2020_Telework <- df_2020 %>% 
  filter(COVIDTelework != 'NIU') %>% 
  mutate(TeleworkCOVID = ifelse(COVIDTelework == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)
## COVIDTelework Breakdown by Employed
TeleworkInd_2020 <- df_2020_Telework %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgTelework = 1 - weighted.mean(TeleworkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))


# Factor industry
TeleworkInd_2020$Industry <- factor(TeleworkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))


p10 <- ggplot(data = TeleworkInd_2020, aes(x = Employed, y = AvgTelework, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgTelework ), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  facet_wrap(~ Industry) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  theme(legend.position="none") +
  labs(title = "Persons still employed in the Retail and Arts, Leisure and Hospitality industries \nreport the highest percentages of being able to telework ", 
       subtitle = "", 
       x = "Employed", y = "% Reporting Able to Telework", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 

# Telework

df_2020_Telework <- df_2020 %>% 
  filter(COVIDTelework != 'NIU') %>% 
  mutate(TeleworkCOVID = ifelse(COVIDTelework == 'No', 0, 1)) %>% 
  select(-COVIDUnableToWork)
## COVIDTelework Breakdown by Employed
TeleworkInd_2020 <- df_2020_Telework %>% 
  group_by(Industry, Employed) %>% 
  summarize(AvgTelework = 1 - weighted.mean(TeleworkCOVID, w = WTFINL)) %>% 
  mutate(Employed = ifelse(Employed == 0, 'No', 'Yes'))


# Factor telework
TeleworkInd_2020$Industry <- factor(TeleworkInd_2020$Industry, levels =c("Arts, Leisure and Hospitality",  "Information", "Manufacturing", "Retail"  ))


p10 <- ggplot(data = TeleworkInd_2020, aes(x = Employed, y = AvgTelework, fill = Industry)) +
  geom_point(color="#011f4b", size=2, alpha=0.6)+
  geom_segment(aes(x=Employed, xend=Employed, y=0, yend=AvgTelework ), color =  "#03396c") +
  #geom_bar(stat = 'identity', position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values=c("#b3cde0", "#6497b1", "#005b96", "#03396c", "#011f4b"))+
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  guides(fill = guide_legend(reverse=TRUE)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(legend.key=element_blank())+
  facet_wrap(~ Industry) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  theme_clean() +
  theme(legend.position="none") +
  labs(title = "Persons still employed in the Retail and Arts, Leisure and Hospitality industries \nreport the highest percentages of being able to telework ", 
       subtitle = "", 
       x = "Employed", y = "% Reporting Able to Telework", 
       tag = , caption = "Source: IPUMS CPS Monthly Data") 




```

```{r, echo=FALSE}
# Print plots

p9
```




\

- `Retail` reported the highest percentage of being unable to work due to COVID-19 .

- While `Arts, Leisure and Hospitality` continues to be the most affected industry among the four looked at, the percentage of persons responding 'no' to being unable to work due to the pandemic was lowest in the  `Information` industry. 

\


```{r, echo=FALSE}
# Print plots

p10
```

\


- Persons still employed in the `Retail` and `Arts,Leisure and Hospitality` industries report the highest percentages of being able to telework.

- By comparison, those still employed in `Information`, a higher sub-group than other industries, report the lowest percentage of being able to telework.


\

**Summary of Exploratory Findings**

\

In looking at the data, overall job numbers show a decrease from 2019 to 2020 for all industries during 2020 due to the shutdowns caused by the COVID-19 pandemic. Manufacturing and Retail are similarly represented in the data and show similar trends. Three inflection points present themselves in the data:

1)	The States shutting down in March due to COVID-19.
2)	The partial reopening that some States undertook in May.
3)	The pause on reopening that occurred in July.

Since the staggered nationwide shutdowns began in March, we used data from April- December 2020 to get a more accurate effect on industry employment rates. The results show that Arts, Leisure and Hospitality is the only industry more affected than Retail in terms of unemployment rates during the pandemic.

The partial reopening helped industries recover from their April unemployment highs but the July pause in reopening stalled recovery for all analyzed industries. When compared to Retail,  Manufacturing has a lower unemployment peak and a less prolonged elevation in unemployment rates .The Information industry's improvement in unemployment rates gained momentum in July, outpacing Retail’s recovery by November 2020.

When looking at those who were unable to work because their employer closed or lost business due to the coronavirus pandemic, persons employed in `Retail` and `Manufacturing` occupations were among the most likely to have been unable to work due to the pandemic during the period of April-December 2020. 

\

### Research Question: *How has retail fared relative to other industries?*

\


- Based on the findings in the exploratory analysis, we will be looking at the effect of industry type on the probability of being employed. Additionally, we will be looking at what effect resulted on the probability of being employed when we interact both `Month` and `Industry` in Retail and the other industries. We will also measure the effect `COVIDUnableToWork` and `CovidTelework` had on the same. We will be confirming if second to Arts, Leisure and Hospitality, Retail was the second most affected industry by COVID-19. Additionally, due to the seasonal nature of `Retail` we look to confirm our findings that `Months` played a role in the probability of being employed. 


\


- As a baseline, we run a model looking solely at the effect the type of industry has on the probability being employed.

\

$$ Employed_i= \beta_0 + \beta_1Industry_i    $$
\

- Our second model looks at the effect of month on the industries, particularly the months associated with the initial COVID-19 shutdown (April), the second shutdown (July), and the peak `Retail` season
of November and December.

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2Month_i\times Industry_i   $$
\

- The third model measures the effect of the interaction between `Industry` and `COVIDUnableToWork` and its effect on the probability of being employed. 

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2COVIDUnableToWork_i\times Industry_i   $$

\







```{r, echo=FALSE, include=FALSE}
# Base model

Model_1 <- lm_robust(Employed ~ Industry , data = df_2020, weights = WTFINL)


# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_1, matchCoefs(Model_1, "Industry"))


# Factor Month
df_2020$MONTH = factor(df_2020$MONTH)

# Industry * Month
Model_2 <- lm_robust(Employed ~ Industry + Industry*MONTH , data = df_2020, weights = WTFINL)


# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_2, matchCoefs(Model_2, "Industry"))


```



\


\

```{r, echo=FALSE, include=FALSE}

## COVID MODELS

# Model with interaction Industry * COVIDUNABLE TO WORK
Model_3 <- lm_robust(Employed ~ Industry + Industry*COVIDUnableToWork, data = df_2020_COVID, weights = WTFINL)


# Perform hypothesis test that all coefficients are equal to 0
linearHypothesis(Model_3, matchCoefs(Model_3, "Industry"))

# Factor Month
df_2020_Telework$COVIDTelework = factor(df_2020_Telework$COVIDTelework)





```

\


##### Model Results:  Industry, Industry:Month, Industry:COVIDUnableToWork


\


```{r, echo=FALSE}
# Compare Models 1, 2

export_summs(Model_1, Model_2, Model_3,
             model.names = c("Industry", "Industry:Month", "Industry:COVIDUnableToWork"),
             coefs = c("Information" = "IndustryInformation",
                       "Manufacturing" = "IndustryManufacturing",
                       "Retail" = "IndustryRetail",
                       "April" = "MONTH4",
                       "July" = "MONTH7",
                       "November" = "MONTH11",
                       "December" = "MONTH12",
                       "Information:April" = "IndustryInformation:MONTH4",
                       "Manufacturing:April" = "IndustryManufacturing:MONTH4",
                       "Retail:April" = "IndustryRetail:MONTH4",
                       "Information:July" = "IndustryInformation:MONTH7",
                       "Manufacturing:July" = "IndustryManufacturing:MONTH7",
                       "Retail:July" = "IndustryRetail:MONTH7",
                       "Information:November" = "IndustryInformation:MONTH11",
                       "Manufacturing:November" = "IndustryManufacturing:MONTH11",
                       "Retail:November" = "IndustryRetail:MONTH11",
                       "Information:December" = "IndustryInformation:MONTH12",
                       "Manufacturing:December" = "IndustryManufacturing:MONTH12",
                       "Retail:December" = "IndustryRetail:MONTH12",
                       "Retail:COVID-AbleToWork" = "IndustryRetail:COVIDUnableToWorkYes",
                       "Manufacturing:COVID-AbleToWork" = "IndustryManufacturing:COVIDUnableToWorkYes",
                       "Information:COVID-AbleToWork" = "IndustryInformation:COVIDUnableToWorkYes"),
                scale = TRUE, robust = TRUE)

```
\

**Key Findings**

\

$$ Employed_i= \beta_0 + \beta_1Industry_i $$

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2Month_i\times Industry_i   $$

\

$$ Employed_i= \beta_0 + \beta_1Industry_i + \beta_2COVIDUnableToWork_i\times Industry_i   $$

\


- **Industry**

  - Being in the `Retail` industry results in a 0.10 higher probability of being employed as opposed to the reference industry, `Arts, Leisure and Entertainment`.

  - Being in the `Information` and `Manufacturing` industries result in a 0.12 and 0.11 higher probability (respectively) of being employed as opposed to the reference industry, `Arts, Leisure and Entertainment.`

- **Industry:Month**

  - Being in the month of April 2020 results in a 0.32 less probability of being employed as opposed to being the reference month of January, 2020.

  - Being in the month of July 2020 results in a 0.18 less probability of being employed as opposed to being the reference month of January, 2020

  - Being in the month of November 2020 results in a 0.09 less probability of being employed as opposed to being the reference month of January, 2020

  - Being in the month of December 2020 results in a 0.10 less probability of being employed as opposed to being in the reference month of January, 2020

  - Being in the month of April 2020 and in the `Retail` industry results in a 0.19 higher probability of being employed as opposed to being in the reference month of January, 2020.
  
  - Being in the month of April 2020 and in both the `Information` and `Manufacturing` industries results in a 0.19 higher probability of being employed as opposed to being in the reference month of January, 2020.and the reference industry,  `Arts, Leisure and Entertainment`.

  - Being in the month of July 2020 and in the `Retail` or the `Manufacturing` industries result in a 0.13 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`.

  - Being in the month of July 2020 and in the `Information` industry results in a 0.08 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry, `Arts, Leisure and Entertainment`.

  - Being in the month of November 2020 and in the `Retail` industry results in a 0.08 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`.

  - Being in the month of November 2020 and in the `Information` industry results in a 0.03 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`.

  - Being in the month of November 2020 and in the `Manufacturing` industry results in a 0.07 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`.

  - Being in the month of December 2020 and in both the `Retail` and `Manufacturing` industries results in a 0.09 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`

  - Being in the month of November 2020 and in the `Information` industry results in a 0.06 higher probability of being employed as opposed to being in both the reference month of January, 2020 and the reference industry,  `Arts, Leisure and Entertainment`.
  
- **Industry:Covid-AbleToWork**

  - Reporting being able to work during the COVID-19 pandemic and being in the `Retail` industry results in a 0.06 higher probability of being employed as opposed to responding as unable to work and being in the reference industry,  `Arts, Leisure and Entertainment`.
  
  - Reporting being able to work during the COVID-19 pandemic and being in the `Manufacturing` industry results in a 0.04 higher probability of being employed as opposed to responding as unable to work and being in the reference industry,  `Arts, Leisure and Entertainment`.
  
  - Reporting being able to work during the COVID-19 pandemic and being in the `Information` industry results in a 0.01 higher probability of being employed as opposed to responding as unable to work and being in the reference industry,  `Arts, Leisure and Entertainment`(This result is statistically insignificant). 


\

**Summary of Findings**

When comparing the Retail to the Arts, Leisure and Entertainment, there is a 10 percentage point higher probability of being employed during the COVID-19 pandemic. Compared to Retail, being in the Manufacturing and Information industries improves the probability of being employed by 1 and 2 percentage points, respectively. 

Our exploratory analysis revealed that April and July were two critical months in 2020 that pointed at increased unemployment rates due to either States shutting down or pausing their reopening. The analysis confirms this. The probability of being employed decreases by 32 percentage points in April and by 18 percentage points in July. Compared to Retail, there is a 3-percentage point higher probability of being employed in both the Manufacturing and Information industries. 

Interestingly, when States paused reopening plans in July, being employed in Retail results in a 5-percentage point higher probability of being employed when compared to the Information industry and a 13-percentage point higher probability when compared to Arts, Leisure and Hospitality. 

As November and December are considered peak season for Retail, we compared its performance to the other industries for those months. When compared to the Information industry, Retail enjoys a 5-percentage point advantage in the probability of being employed in November and a 3-percentage point advantage in December. When compared to Arts, Leisure and Hospitality, the same December probability advantage for Retail increases to 9-percentage points. 

Finally, in looking at the special COVID-19 question of  "At any time in the last 4 weeks, were you unable to work because your employer closed or lost business due to the coronavirus pandemic?" reveals that answering yes to this question and being in the Retail industry results in a 6-percentage point higher probability of being employed when compared to the Information industry and 6-percentage point advantage over the Arts, Leisure and Hospitality industry. 


\

































